---
title: "SSL Future Proof Project Update - Jan 2023"
output: html_document
---

```{r setup, include=FALSE, fig.align = 'center'}

knitr::opts_chunk$set(echo = FALSE, eval = T, message = F, warning = F, fig.align = 'center')

library(here) #v1.0.1
library(dplyr) #v1.0.5
library(nimble) #v0.12.1
library(coda)
library(ggh4x)
library(readr)
library(knitr)
library(Hmisc)
library(stringr)

source(here::here('scripts', 'PlotTheme.R'))

```

```{r scenarios and convergence}

sc <- c(1:45)

scenarios <- data.frame(scenario = c(sc),
                        survey_freq = c(rep(1,15), rep(2,15), rep(3,15)),
                        detection = c(rep(c('Low', 'Low', 'Low', 'Low', 'Low',
                                            'Med', 'Med', 'Med', 'Med', 'Med', 
                                            'High', 'High', 'High', 'High', 'High'), 3)),
                        N = c(rep(c(50,75,100,150,200), 9)))


#for figure legends
phi_pars <- c('phi[P]','phi[1]', 'phi[2]', 'phi[J]', 'phi[A]')
p_pars <- c('p[1]', 'p[2]', 'p[J]', 'p[A]') 
freq_labels <- c('1' = 'Annual', '2' = 'Biennial', '3' = 'Triennial')
occ_labels <- c('10' = '10 yrs', '20' = '20 yrs')

```


```{r cv}

#load in results for 10-yr study horizon
cv.files7 <- list.files(path = here::here('results', 'simpleAge', 'sims'), 
                    pattern = 'cv7', all.files = FALSE, full.names = F)

cv.files8 <- list.files(path = here::here('results', 'simpleAge', 'sims'), 
                    pattern = 'cv8', all.files = FALSE, full.names = F)

cv.files9 <- list.files(path = here::here('results', 'simpleAge', 'sims'), 
                    pattern = 'cv9', all.files = FALSE, full.names = F)

cv.files10 <- list.files(path = here::here('results', 'simpleAge', 'sims'), 
                    pattern = 'cv10', all.files = FALSE, full.names = F)

cv.files11 <- list.files(path = here::here('results', 'simpleAge', 'sims'), 
                    pattern = 'cv11', all.files = FALSE, full.names = F)

cv.files12 <- list.files(path = here::here('results', 'simpleAge', 'sims'), 
                    pattern = 'cv12', all.files = FALSE, full.names = F)

cv.files <- c(cv.files7, cv.files8, cv.files9, cv.files10, cv.files11, cv.files12)


cv <- data.frame()
for (f in 1:length(cv.files)) {
temp.cv <- data.frame(readRDS(file = here::here('results', 'simpleAge', 'sims',
                                                  cv.files[f])))
#use gsub on cv files rather than use different list to make sure they match
rhats <- data.frame(rhat = readRDS(file = here::here('results', 'simpleAge', 'sims', 
                                                     gsub('cv', 'median', cv.files[f])))[,'max_rhat'])

temp.cv$rhat <- as.numeric(rhats$rhat)
temp <- temp.cv %>% 
  reshape2::melt(id.vars = 'rhat') %>%
  transform(scenario = parse_number(gsub('S.', '', cv.files[f]))) %>%
  transform(value = ifelse(is.nan(value), 0, value))

cv <- rbind(cv, temp)

} #rds file

cv <- cv %>%
  filter(rhat < 1.1) 

mean.cv <- data.frame(cv) %>%
  group_by(variable, scenario) %>%
  dplyr::summarize(value = mean(value), .groups = 'keep') %>%
  merge(scenarios, by = 'scenario', all = F) %>%
  group_by(variable, detection, N, survey_freq) %>%
  dplyr::summarize(value = mean(value), .groups = 'keep') %>%
  transform(variable = factor(variable, levels = c('phiP', 'phi1', 'phi2', 'phiJ', 'phiA',
                            'p1', 'p2', 'pJ', 'pA'))) %>%
  transform(detection = factor(detection, levels = c('Low', 'Med', 'High'), labels = c('Low', 'Medium', 'High'))) %>%
  transform(survey_frequency = 'Branding frequency & study length', samp_size = 'Sample size') %>%
  transform(length = 10)

mean.cv.thresh <- data.frame(cv) %>%
  transform(value_cat = value<0.125) %>%
  group_by(variable, scenario) %>%
  dplyr::summarize(value_cat = mean(value_cat), .groups = 'keep') %>%
  merge(scenarios, by = 'scenario', all = F) %>%
  group_by(variable, detection, N, survey_freq) %>%
  dplyr::summarize(value_cat = mean(value_cat), .groups = 'keep') %>%
  transform(variable = factor(variable, levels = c('phiP', 'phi1', 'phi2', 'phiJ', 'phiA'))) %>%
  transform(detection = factor(detection, levels = c('Low', 'Med', 'High'), labels = c('Low', 'Medium', 'High'))) %>%
  transform(survey_frequency = 'Branding frequency & study length', samp_size = 'Sample size') %>%
  transform(length = 10)

#load in results from 20-yr study and merge
cv.files10 <- list.files(path = here::here('results', 'simpleAge', 'sims', 'prePop'), 
                    pattern = 'cv10', all.files = FALSE, full.names = F)

cv.files11 <- list.files(path = here::here('results', 'simpleAge', 'sims', 'prePop'), 
                    pattern = 'cv11', all.files = FALSE, full.names = F)

cv.files14 <- list.files(path = here::here('results', 'simpleAge', 'sims', 'prePop'), 
                    pattern = 'cv14', all.files = FALSE, full.names = F)

cv.files15 <- list.files(path = here::here('results', 'simpleAge', 'sims', 'prePop'), 
                    pattern = 'cv15', all.files = FALSE, full.names = F)

cv.files18 <- list.files(path = here::here('results', 'simpleAge', 'sims', 'prePop'), 
                    pattern = 'cv18', all.files = FALSE, full.names = F)

cv.files19 <- list.files(path = here::here('results', 'simpleAge', 'sims', 'prePop'), 
                    pattern = 'cv19', all.files = FALSE, full.names = F)

cv.files22 <- list.files(path = here::here('results', 'simpleAge', 'sims', 'prePop'),
                         pattern = 'cv22', all.files = FALSE, full.names = F)

cv.files23 <- list.files(path = here::here('results', 'simpleAge', 'sims', 'prePop'), 
                    pattern = 'cv23', all.files = FALSE, full.names = F)

cv.files26 <- list.files(path = here::here('results', 'simpleAge', 'sims', 'prePop'), 
                    pattern = 'cv26', all.files = FALSE, full.names = F)

cv.files27 <- list.files(path = here::here('results', 'simpleAge', 'sims', 'prePop'), 
                    pattern = 'cv27', all.files = FALSE, full.names = F)

cv.files30 <- list.files(path = here::here('results', 'simpleAge', 'sims', 'prePop'), 
                    pattern = 'cv30', all.files = FALSE, full.names = F)

cv.files31 <- list.files(path = here::here('results', 'simpleAge', 'sims', 'prePop'), 
                    pattern = 'cv31', all.files = FALSE, full.names = F)


cv.files <- c(cv.files10, cv.files11, cv.files14, cv.files15, cv.files18, cv.files19,
              cv.files22, cv.files23, cv.files26, cv.files27, cv.files30, cv.files31)

cv.20 <- data.frame()
for (f in 1:length(cv.files)) {
temp.cv <- data.frame(readRDS(file = here::here('results', 'simpleAge', 'sims', 'prePop',
                                                  cv.files[f])))
#use gsub on cv files rather than use different list to make sure they match
rhats <- data.frame(rhat = readRDS(file = here::here('results', 'simpleAge', 'sims', 'prePop',
                                                     gsub('cv', 'median', cv.files[f])))[,'max_rhat'])

temp.cv$rhat <- as.numeric(rhats$rhat)
temp <- temp.cv %>% 
  reshape2::melt(id.vars = 'rhat') %>%
  transform(scenario = parse_number(gsub('S.', '', cv.files[f]))) %>%
  transform(value = ifelse(is.nan(value), 0, value))

cv.20 <- rbind(cv.20, temp)

} #rds file

cv.20 <- cv.20 %>%
  filter(rhat < 1.1) #%>%
  # transform(length = 20) #%>%
  # bind_rows(cv)

mean.cv.20 <- data.frame(cv.20) %>%
  group_by(variable, scenario) %>%
  dplyr::summarize(value = mean(value), .groups = 'keep') %>%
  merge(scenarios, by = 'scenario', all = F) %>%
  group_by(variable, detection, N, survey_freq) %>%
  dplyr::summarize(value = mean(value), .groups = 'keep') %>%
  transform(variable = factor(variable, levels = c('phiP', 'phi1', 'phi2', 'phiJ', 'phiA',
                            'p1', 'p2', 'pJ', 'pA'))) %>%
  transform(detection = factor(detection, levels = c('Low', 'Med', 'High'), labels = c('Low', 'Medium', 'High'))) %>%
  transform(survey_frequency = 'Branding frequency & study length', samp_size = 'Sample size') %>%
  transform(length = 20) %>%
  bind_rows(mean.cv)

### figures

## continuous raw all facets -- for appendix
plot_dat <- mean.cv.20 %>% filter(grepl('phi', variable)) %>%
  transform(variable = factor(variable, levels = c('phiP', 'phi1', 'phi2', 'phiJ', 'phiA'), 
                              labels = phi_pars))

cv_heat_cont <- ggplot(plot_dat, aes(x = factor(detection), y = variable, fill = value)) +
  geom_tile(color = 'grey50') +
  xlab('Detection') +
  ylab('Parameter') +
  facet_nested(samp_size + N ~ survey_frequency + survey_freq + length, drop = T, 
             labeller = labeller(survey_freq = freq_labels, length = occ_labels),
             scales = 'free_x') +
 scale_fill_gradient2(name = "cv",
                      mid = "#FFFFFF", low = "#012345", high = "#012345", midpoint = 0) +
  theme_bw() +
  theme(legend.position = 'right',
             strip.background = element_rect(fill = "#EEEEEE", color = "#FFFFFF"),
        legend.text = element_text(size = 8, vjust = 0.5),
        legend.title = element_text(vjust = 0.8, size = 10)) +
    scale_y_discrete("", labels = parse(text = c(eval(expression(levels(plot_dat$variable))))))

## continuous raw dotplots
dotplot_dat_mean <- cv.20 %>% transform(length = 20) %>% 
  bind_rows(cv %>% transform(length = 10)) %>%
  merge(scenarios, by = 'scenario', all = F) %>%
  group_by(length, variable, detection, survey_freq, N) %>%
    dplyr::summarize(value = mean(value), .groups = 'keep') %>%
  transform(detection = factor(detection, levels = c('Low', 'Med', 'High'), labels = c('Low', 'Medium', 'High'))) %>%
  #filter out detection params
  filter(!is.na(variable))

dotplot_dat_up <- cv.20 %>% transform(length = 20) %>% 
  bind_rows(cv %>% transform(length = 10)) %>%
  merge(scenarios, by = 'scenario', all = F) %>%
  group_by(length, variable, detection, survey_freq, N) %>%
  dplyr::summarize(upper = quantile(value, probs = 0.975)) %>%
  transform(detection = factor(detection, levels = c('Low', 'Med', 'High'), labels = c('Low', 'Medium', 'High'))) %>%
  #filter out detection params
  filter(!is.na(variable))

dotplot_dat_low <- cv.20 %>% transform(length = 20) %>% 
  bind_rows(cv %>% transform(length = 10)) %>%
  merge(scenarios, by = 'scenario', all = F) %>%
  group_by(length, variable, detection, survey_freq, N) %>%
  dplyr::summarize(lower = quantile(value, probs = 0.025)) %>%
  transform(detection = factor(detection, levels = c('Low', 'Med', 'High'), labels = c('Low', 'Medium', 'High'))) %>%
  #filter out detection params
  filter(!is.na(variable))

dotplot_dat <- dotplot_dat_mean %>%
  merge(dotplot_dat_low, by = c('length', 'variable', 'detection', 'survey_freq', 'N')) %>%
  merge(dotplot_dat_up, by = c('length', 'variable', 'detection', 'survey_freq', 'N')) %>%
  transform(variable = factor(variable, levels = c('phiP', 'phi1', 'phi2', 'phiJ', 'phiA'), 
                              labels = c('Pup', 'Age-1', 'Age-2', 'Juvenile', 'Adult'))) %>%
  transform(survey_freq = factor(survey_freq, levels = c(1,2,3),
                                 labels = c('Annual', 'Biennial', 'Triennial'))) %>%
  transform(survey_frequency = 'Branding frequency', samp_size = 'Sample size',
            det = 'Detection')

#no study horizon
cv_cont_dot <- ggplot(dotplot_dat %>% filter(N %in% c(50, 100, 200) & length == 20),
       aes(x = survey_freq, y = value, col = factor(variable), group = variable)) +
  geom_point() + geom_line() +
  geom_errorbar(aes(ymin = lower, ymax = upper, x = survey_freq), width = 0.1) +
  xlab('Branding frequency') + ylab('CV') +
  facet_nested(det + detection ~ samp_size + N, drop = T,
             labeller = labeller(survey_freq = freq_labels, length = occ_labels),
             scales = 'free_x') +
  theme_bw() +
  plot_theme(legend.position = 'top',
             plot.subtitle = element_text(size = 10, hjust = 0.5, vjust = 1)) +
  scale_color_manual(values = rainbow2, name = 'Survival rate')


##threshold approach

#heatmaps
mean.cv.thresh.20 <- data.frame(cv %>% transform(length = 10)) %>% 
  bind_rows(cv.20 %>% transform(length = 20)) %>%
  transform(value_cat = value<0.125) %>%
  group_by(variable, scenario, length) %>%
  dplyr::summarize(value_cat = mean(value_cat), .groups = 'keep') %>%
  merge(scenarios, by = 'scenario', all = F) %>%
  group_by(variable, detection, N, survey_freq, length) %>%
  dplyr::summarize(value_cat = mean(value_cat), .groups = 'keep') %>%
  transform(variable = factor(variable, levels = c('phiP', 'phi1', 'phi2', 'phiJ', 'phiA'))) %>%
  transform(detection = factor(detection, levels = c('Low', 'Med', 'High'), 
                               labels = c('Low', 'Medium', 'High'))) %>%
  transform(survey_frequency = 'Branding frequency & study length', samp_size = 'Sample size')

plot_dat <- mean.cv.thresh.20 %>% filter(grepl('phi', variable)) %>%
  transform(variable = factor(variable, levels = c('phiP', 'phi1', 'phi2', 'phiJ', 'phiA'), 
                              labels = phi_pars))

cv_thresh_heat <- ggplot(plot_dat, aes(x = factor(detection), y = variable, fill = value_cat)) +
  geom_tile(color = 'grey50') +
  xlab('Detection') + ylab('Parameter') +
  facet_nested(samp_size + N ~ survey_frequency + survey_freq + length, drop = T, 
             labeller = labeller(survey_freq = freq_labels, length = occ_labels),
             scales = 'free_x') +
scale_fill_gradient2(name = "Probability of \n cv < 0.125",
                  mid = "#FFFFFF", high = "deepskyblue1", low = "red2", midpoint = 0.8) +
  theme_bw() +
  theme(legend.position = 'right',
             strip.background = element_rect(fill = "#EEEEEE", color = "#FFFFFF"),
        legend.text = element_text(size = 8, vjust = 0.5),
        legend.title = element_text(vjust = 0.8, size = 10)) +
    scale_y_discrete("", labels = parse(text = c(eval(expression(levels(plot_dat$variable))))))


#dotplot threshold
dotplot_dat <- cv.20 %>% transform(length = 20) %>% 
  bind_rows(cv %>% transform(length = 10)) %>%
  merge(scenarios, by = 'scenario', all = F) %>%
  transform(value_cat = value<0.125) %>%
  group_by(length, variable, detection, survey_freq, N) %>%
  dplyr::summarize(value_cat = mean(value_cat), .groups = 'keep') %>%
  transform(detection = factor(detection, levels = c('Low', 'Med', 'High'), labels = c('Low', 'Medium', 'High'))) %>%
  #filter out detection params
  filter(!is.na(variable)) %>%
  transform(variable = factor(variable, levels = c('phiP', 'phi1', 'phi2', 'phiJ', 'phiA'), 
                              labels = c('Pup', 'Age-1', 'Age-2', 'Juvenile', 'Adult'))) %>%
  transform(survey_freq = factor(survey_freq, levels = c(1,2,3),
                                 labels = c('Annual', 'Biennial', 'Triennial'))) %>%
  transform(survey_frequency = 'Branding frequency', samp_size = 'Sample size',
            det = 'Detection')


#no study horizon; no quantile error bars on yes/no 
cv_thresh_dot <- ggplot(dotplot_dat %>% filter(N %in% c(50, 100, 200) & length == 20),
       aes(x = survey_freq, y = value_cat, col = factor(variable), group = variable)) +
  geom_point() + geom_line() +
  xlab('Branding frequency') + ylab('Probability of cv < 0.125') +
  facet_nested(det + detection ~ samp_size + N, drop = T,
             labeller = labeller(survey_freq = freq_labels, length = occ_labels),
             scales = 'free_x') +
  theme_bw() +
  plot_theme(legend.position = 'top',
             plot.subtitle = element_text(size = 10, hjust = 0.5, vjust = 1)) +
  scale_color_manual(values = rainbow2, name = 'Survival rate')



```


```{r detecting change}

#load in results from 10-yr
cri.scenarios <- c('dec_little', 'dec_big')
append <- c(5,10)

cri.all <- data.frame()

for (a in 1:length(append)) { #files for period length

for (j in 1:length(cri.scenarios)) {
  
cri.files <- list.files(path = here::here('results', 'simpleAge', 'sims'), 
                    pattern = paste0(cri.scenarios[j],append[a]), all.files = FALSE, full.names = F)

cri <- data.frame()
for (f in 1:length(cri.files)) {
temp.cri <- data.frame(readRDS(file = here::here('results', 'simpleAge', 'sims',
                                                  cri.files[f])))
#use gsub on cri files rather than use different list to make sure they match
rhats <- data.frame(rhat = readRDS(file = here::here('results', 'simpleAge', 'sims',
                                                     gsub('median_', 'median', gsub(paste0(cri.scenarios[j], append[a]), 'median', cri.files[f]))))[,'max_rhat'])

temp.cri$rhat <- as.numeric(rhats$rhat)
temp <- temp.cri %>% 
  reshape2::melt(id.vars = 'rhat') %>%
  transform(scenario = parse_number(gsub('S.', '', cri.files[f]))) %>%
  transform(run = str_sub(cri.files[f], -5,-5)) %>%
  transform(length = append[a]) %>%
  # transform(length = ifelse(grepl('little5', cri.files[f]) | grepl('big5', cri.files[f]), 5, 10)) %>%
  transform(threshold = cri.scenarios[j])

cri <- rbind(cri, temp)

} #appending length of year period

cri.all <- rbind(cri, cri.all)

} #rds file
} #j


mean.cri <- data.frame(cri.all) %>%
  filter(!is.na(value)) %>%
    merge(scenarios, by = 'scenario', all = F) %>%
  group_by(threshold, variable, detection, N, survey_freq, length) %>%
  dplyr::summarize(value = mean(value), .groups = 'keep') %>%
  transform(variable = factor(variable, levels = c('phiP', 'phi1', 'phi2', 'phiJ', 'phiA',
                            'p1', 'p2', 'pJ', 'pA'))) %>%
  transform(detection = factor(detection, levels = c('Low', 'Med', 'High'))) %>%
  transform(survey_frequency = 'Branding frequency & study length', samp_size = 'Sample size') 

#load in results from 20-yr
cri.all.20 <- data.frame()

for (a in 1:length(append)) { #files for period length

for (j in 1:length(cri.scenarios)) {
  
cri.files <- list.files(path = here::here('results', 'simpleAge', 'sims', 'prePop'), 
                    pattern = paste0(cri.scenarios[j],append[a]), all.files = FALSE, full.names = F)

cri <- data.frame()
for (f in 1:length(cri.files)) {
temp.cri <- data.frame(readRDS(file = here::here('results', 'simpleAge', 'sims', 'prePop',
                                                  cri.files[f])))
#use gsub on cri files rather than use different list to make sure they match
rhats <- data.frame(rhat = readRDS(file = here::here('results', 'simpleAge', 'sims', 'prePop', 
                                                     gsub('median_', 'median', gsub(paste0(cri.scenarios[j], append[a]), 'median', cri.files[f]))))[,'max_rhat'])

temp.cri$rhat <- as.numeric(rhats$rhat)
temp <- temp.cri %>% 
  reshape2::melt(id.vars = 'rhat') %>%
  transform(scenario = parse_number(gsub('S.', '', cri.files[f]))) %>%
  transform(run = str_sub(cri.files[f], -6,-5)) %>%
  transform(length = append[a]+10) %>%
  # transform(length = ifelse(grepl('little5', cri.files[f]) | grepl('big5', cri.files[f]), 5, 10)) %>%
  transform(threshold = cri.scenarios[j])

cri <- rbind(cri, temp)

} #appending length of year period

cri.all.20 <- rbind(cri, cri.all.20)

} #rds file
} #j

mean.cri.20 <- data.frame(cri.all.20) %>%
  #are these just all the unfinished/unfilled sim dataframe rows? 
  filter(!is.na(value)) %>%
    merge(scenarios, by = 'scenario', all = F) %>%
  group_by(threshold, variable, detection, N, survey_freq, length) %>%
  dplyr::summarize(value = mean(value), .groups = 'keep') %>%
  transform(variable = factor(variable, levels = c('phiP', 'phi1', 'phi2', 'phiJ', 'phiA'))) %>%
  transform(detection = factor(detection, levels = c('Low', 'Med', 'High'))) %>%
  bind_rows(mean.cri)

#plots

#all facets
plot_dat <- mean.cri.20 %>% filter(grepl('phi', variable)) %>%
  transform(variable = factor(variable, levels = c('phiP', 'phi1', 'phi2', 'phiJ', 'phiA'), 
                              labels = phi_pars)) %>%
  transform(value_cat = ifelse(value < 0.5, "<0.5",
                               ifelse(value >=0.5 & value < 0.8, '0.5-0.8', '>0.8'))) %>%
  transform(value_cat = factor(value_cat, levels = c('<0.5', '0.5-0.8', '>0.8'))) %>%
  transform(threshold = factor(threshold, levels = c('dec_little', 'dec_big'),
                               labels = c('Small decrease', 'Big decrease'))) %>%
  transform(detection = factor(detection, levels = c('Low', 'Med', 'High'),
                               labels = c('Low', 'Medium', 'High'))) %>%
  transform(survey_frequency = 'Branding frequency & change size', 
            samp_size = 'Sample size & study length')

#all facets continuous
cri_all_cont_heat <- ggplot(plot_dat %>% filter(length %in% c(10,20)),
       aes(x = factor(detection), y = variable, fill = value)) +
        # aes(x = factor(detection), y = variable, fill = value_cat)) +
  geom_tile(color = 'grey50') +
  xlab('Detection') +
  ylab('Parameter') +
  facet_nested(samp_size + N + length ~ survey_frequency + survey_freq + threshold, drop = T,
             labeller = labeller(survey_freq = freq_labels, length = occ_labels),
             scales = 'free_x') +
scale_fill_gradient2(name = "Probability of detecting change",
                  mid = "#FFFFFF", high = "deepskyblue1", low = "red2", midpoint = 0.8) +
  theme_bw() +
  theme(legend.position = 'top',
             strip.background = element_rect(fill = "#EEEEEE", color = "#FFFFFF")) +
    scale_y_discrete("", labels = parse(text = c(eval(expression(levels(plot_dat$variable))))))

#all facets categorical
cri_all_cat_heat <- ggplot(plot_dat %>% filter(length %in% c(10,20)),
       aes(x = factor(detection), y = variable, fill = value_cat)) +
  geom_tile(color = 'grey50') +
  xlab('Detection') +
  ylab('Parameter') +
  facet_nested(samp_size + N + length ~ survey_frequency + survey_freq + threshold, drop = T,
             labeller = labeller(survey_freq = freq_labels, length = occ_labels),
             scales = 'free_x') +
  theme_bw() +
  theme(legend.position = 'top',
             strip.background = element_rect(fill = "#EEEEEE", color = "#FFFFFF")) +
    scale_y_discrete("", labels = parse(text = c(eval(expression(levels(plot_dat$variable)))))) +
  scale_fill_manual(name = "Probability of detecting change in survival",
                    # values = c('firebrick3', 'orange', 'steelblue2'))
                    values = c('salmon', 'lightgoldenrod2', 'lightskyblue2'))

#fewer facets
 plot_dat <- mean.cri.20 %>% filter(grepl('phi', variable) & length %in% c(10,20)) %>%
  transform(variable = factor(variable, levels = c('phiP', 'phi1', 'phi2', 'phiJ', 'phiA'), 
                              labels = phi_pars)) %>%
  transform(value_cat = ifelse(value < 0.5, "<0.5",
                               ifelse(value >=0.5 & value < 0.8, '0.5-0.8', '>0.8'))) %>%
  transform(value_cat = factor(value_cat, levels = c('<0.5', '0.5-0.8', '>0.8'))) %>%
  transform(threshold = factor(threshold, levels = c('dec_little', 'dec_big', 'inc_little', 'inc_big'),
                               labels = c('Small decrease', 'Big decrease', 'Small increase', 'Big increase'))) %>%
  transform(survey_frequency = 'Branding frequency & study length', samp_size = 'Sample size')
 
#big decrease in survival 
cri_big_heat <- ggplot(plot_dat %>% filter(threshold %in% c('Big decrease')), 
       aes(x = factor(detection), y = variable, fill = value)) +
        # aes(x = factor(detection), y = variable, fill = value_cat)) +
  geom_tile(color = 'grey50') +
  xlab('Detection') +
  ylab('Parameter') +
  facet_nested(samp_size + N ~ survey_frequency + survey_freq + length, drop = T, 
             labeller = labeller(survey_freq = freq_labels, length = occ_labels),
             scales = 'free_x') +
scale_fill_gradient2(name = "Probability of detecting change",
                  mid = "#FFFFFF", high = "deepskyblue1", low = "red2", midpoint = 0.8) +
  theme_bw() +
  theme(legend.position = 'top',
             strip.background = element_rect(fill = "#EEEEEE", color = "#FFFFFF")) +
    scale_y_discrete("", labels = parse(text = c(eval(expression(levels(plot_dat$variable)))))) 

#categorical, small decrease
cri_big_cat_heat <- ggplot(plot_dat %>% filter(threshold %in% c('Big decrease')),
       aes(x = factor(detection), y = variable, fill = value_cat), alpha = 0.3) +
  geom_tile(color = 'grey50') +
  xlab('Detection') +
  ylab('Parameter') +
  facet_nested(samp_size + N ~ survey_frequency + survey_freq + length, drop = T, 
             labeller = labeller(survey_freq = freq_labels, length = occ_labels),
             scales = 'free_x') +
  theme_bw() +
  theme(legend.position = 'top',
             strip.background = element_rect(fill = "#EEEEEE", color = "#FFFFFF")) +
    scale_y_discrete("", labels = parse(text = c(eval(expression(levels(plot_dat$variable)))))) +
  scale_fill_manual(name = "Probability of detecting small change",
                    values = c('salmon', 'lightgoldenrod2', 'lightskyblue2'))

#dotplot
dotplot_dat <- cri.all.20 %>% transform(length = 20) %>% 
  bind_rows(cri.all %>% transform(length = 10)) %>%
  merge(scenarios, by = 'scenario', all = F) %>%
  group_by(length, variable, detection, survey_freq, N, threshold) %>%
    dplyr::summarize(value = mean(value, na.rm = T), .groups = 'keep') %>%
  transform(variable = factor(variable, levels = c('phiP', 'phi1', 'phi2', 'phiJ', 'phiA'),
                              labels = c('Pup', 'Age-1', 'Age-2', 'Juvenile', 'Adult'))) %>%
  transform(detection = factor(detection, levels = c('Low', 'Med', 'High'),
                               labels = c('Low', 'Medium', 'High'))) %>%
  transform(survey_freq = factor(survey_freq, levels = c(1,2,3), 
                                 labels = c('Annual', 'Biennial', 'Triennial'))) %>%
  transform(survey_frequency = 'Branding frequency', samp_size = 'Sample size',
            det = 'Detection') %>%
  #filter out detection params
  filter(!is.na(variable))


#no study horizon
cri_dot <- ggplot(dotplot_dat %>% filter(N %in% c(50, 100, 200) & length == 20
                               &  threshold == 'dec_big'),
       aes(x = factor(survey_freq), y = value, col = factor(variable), group = variable)) +
  geom_point() + geom_line() +
  geom_hline(aes(yintercept = 0), linetype = 'dotted') +
  xlab('Branding frequency') + ylab('Probability of detecting large change in survival') +
  facet_nested(det ~ detection ~ samp_size + N, drop = T,
             labeller = labeller(survey_freq = freq_labels, length = occ_labels),
             scales = 'free_x') +
  theme_bw() +
  plot_theme(legend.position = 'top',
             plot.subtitle = element_text(size = 10, hjust = 0.5, vjust = 1)) +
  scale_color_manual(values = rainbow2, name = 'Survival rate')


```

```{r collapsed detecting change}

#mean.cri.20 contains the mean probability (across simulations) where the cv < 0.125.

#values for the scenario with the "least" amount of contributing information
#i.e., triennial sampling, N = 50, 10 years, low detection (scenario 31)

cri_least_info <- mean.cri.20 %>% transform(samp_size = 'Sample size') %>%
  dplyr::select(!survey_frequency) %>%
  filter(detection == 'Low' & length %in% c(10,20) &
           survey_freq == 3 & N %in% c(50, 100, 200)) %>%
  dplyr::select(c(variable, N, value, length, threshold)) %>%
  dplyr::rename(min_val = value)

cri_delta <- mean.cri.20 %>% filter(length %in% c(10,20)) %>%
    filter(N %in% c(50, 100, 200)) %>% 
  merge(cri_least_info, by = c('variable', 'N', 'length', 'threshold')) %>%
  transform(delta = value-min_val) %>%
  transform(det_cat = as.numeric(factor(detection)),
            N_cat = as.numeric(factor(N))) %>%
  # filter(variable == 'phiP') %>%
  filter(variable %in% c('phiP', 'phi1')) %>%
  transform(survey_cat = 4-as.numeric(factor(survey_freq))) %>%
  transform(step = N_cat + survey_cat + det_cat) %>%
  transform(step = ifelse(detection == 'Low' & survey_freq == 3, 0, step))

cri_delta_steps <- cri_delta %>% 
  # filter(N == 50 & variable == 'phiP') %>%
    filter(variable %in% c('phiP', 'phi1')) %>%
  transform(category = ifelse(step == 0, 'Baseline', 
                  ifelse(det_cat > 1 & survey_cat == 1, 'Detection only',
                  ifelse(survey_cat > 1 & det_cat == 1, 'Frequency only', 'Detection & frequency')))) %>%
  transform(step = ifelse(step == 0, 0,
                          ifelse(det_cat == 1 & survey_cat == 1, NA,
                            ifelse(category == 'Detection only' & N_cat == 1 & det_cat == 2 | 
                            category == 'Frequency only' & N_cat == 1 & survey_cat == 2 |
                              survey_cat == 2 & N_cat == 1 & det_cat == 2, 1,
                          ifelse(category == 'Detection only' & N_cat == 1 & det_cat == 3 | 
                            category == 'Frequency only' & N_cat == 1 & survey_cat == 3 |
                              survey_cat == det_cat & N_cat == 1, 2, 
                            ifelse(category == 'Detection only' & N_cat == 2 & det_cat == 2 | 
                            category == 'Frequency only' & N_cat == 2 & survey_cat == 2 |
                              survey_cat == 2 & N_cat == 2 & det_cat == 2, 1,
                          ifelse(category == 'Detection only' & N_cat == 2 & det_cat == 3 | 
                            category == 'Frequency only' & N_cat == 2 & survey_cat == 3 |
                              survey_cat == det_cat & N_cat == 2, 2, 
                            ifelse(category == 'Detection only' & N_cat == 3 & det_cat == 2 | 
                            category == 'Frequency only' & N_cat == 3 & survey_cat == 2 |
                              survey_cat == 2 & N_cat == 3 & det_cat == 2, 1,
                          ifelse(category == 'Detection only' & N_cat == 3 & det_cat == 3 | 
                            category == 'Frequency only' & N_cat == 3 & survey_cat == 3 |
                              survey_cat == det_cat & N_cat == 3, 2, 
                            ifelse(det_cat == 1 & survey_cat == 1, NA, NA)))))))))) 

scenario_fill_cri <- expand.grid(step = 0,
              length = c(10,20),
              threshold = c('dec_little', 'dec_big'),
              N = c(50, 100, 200),
              category = c('Detection only', 'Frequency only', 
                                         'Detection & frequency')) %>%
  merge(cri_delta_steps %>% dplyr::select(length, threshold, min_val, category, variable, N), 
        by = c('category', 'N', 'length', 'threshold')) %>% 
  distinct() %>%
  dplyr::rename(value = min_val)
                            
plot_dat <- cri_delta_steps %>%
  dplyr::select(category, N, length, threshold, step, value, variable) %>%
  bind_rows(scenario_fill_cri) %>%
  filter(category != 'Baseline') %>%
  transform(threshold_name = factor(threshold, levels = c('dec_little', 'dec_big'), 
                                    labels = c('Small decrease', 'Large decrease'))) %>%
  transform(category = factor(category, levels = c('Baseline', 'Detection only', 
                                 'Frequency only', 'Detection & frequency'),
                              labels = c('Baseline', 'Detection only', 
                                 'Branding frequency only', 'Detection & branding frequency'))) 

#probability of detecting change.... 
effort_dot <- ggplot(plot_dat %>% filter(!is.na(step)) %>% filter(variable == 'phiP') %>%
         transform(samp_size = 'Sample size & study length') %>%
         transform(length = factor(length, labels = c('10 yrs', '20 yrs'))), 
       aes(factor(step), value, group = category, col = category)) +
  # geom_point(position = position_dodge(width = 0.1)) +
  geom_point() +
  geom_line() +
  facet_nested(threshold_name~samp_size+N+length) +
  xlab('Stepwise change in survey effort') + ylab('Probability of detecting change in pup survival') +
  theme_bw() +
  theme(legend.position = 'top',
             strip.background = element_rect(fill = "#EEEEEE", color = "#FFFFFF"),
        legend.text = element_text(size = 8, vjust = 0.5),
        legend.title = element_text(vjust = 0.8, size = 10)) +
    scale_color_manual(values = rainbow2[c(1,2,5,6)], name = '')

```


```{r relative bias}

bias.files <- list.files(path = here::here('results', 'simpleAge', 'sims'), 
                    pattern = 'bias', all.files = FALSE, full.names = F)

rel.bias <- data.frame()
for (f in 1:length(bias.files)) {
temp.bias <- data.frame(readRDS(file = here::here('results', 'simpleAge', 'sims',
                                                  bias.files[f])))
#use gsub on bias files rather than use different list to make sure they match
rhats <- data.frame(rhat = readRDS(file = here::here('results', 'simpleAge', 'sims',
                                                     gsub('bias', 'median', bias.files[f])))[,'max_rhat'])

temp.bias$rhat <- as.numeric(rhats$rhat)
temp <- temp.bias %>% 
  reshape2::melt(id.vars = 'rhat') %>%
  transform(scenario = parse_number(gsub('S.', '', bias.files[f]))) %>%
  transform(value = ifelse(is.nan(value), 0, value))

rel.bias <- rbind(rel.bias, temp)

} #rds file

rel.bias <- rel.bias %>%
  filter(rhat < 1.1) %>%
  transform(length = 10)

mean.rel.bias <- data.frame(rel.bias) %>%
  group_by(variable, scenario) %>%
  dplyr::summarize(value = mean(value), .groups = 'keep') %>%
  merge(scenarios, by = 'scenario', all = F) %>%
  group_by(variable, detection, N, survey_freq) %>%
  dplyr::summarize(value = mean(value)*100, .groups = 'keep') %>%
  transform(variable = factor(variable, levels = c('phiP', 'phi1', 'phi2', 'phiJ', 'phiA',
                            'p1', 'p2', 'pJ', 'pA'))) %>%
  transform(detection = factor(detection, levels = c('Low', 'Med', 'High'))) %>%
  transform(survey_frequency = 'Branding frequency & study length', samp_size = 'Sample size') %>%
  transform(length = 10)


bias.files <- list.files(path = here::here('results', 'simpleAge', 'sims', 'prePop'), 
                    pattern = 'bias', all.files = FALSE, full.names = F)

rel.bias.20 <- data.frame()
for (f in 1:length(bias.files)) {
temp.bias <- data.frame(readRDS(file = here::here('results', 'simpleAge', 'sims', 'prePop',
                                                  bias.files[f])))
#use gsub on bias files rather than use different list to make sure they match
rhats <- data.frame(rhat = readRDS(file = here::here('results', 'simpleAge', 'sims', 'prePop',
                                                     gsub('bias', 'median', bias.files[f])))[,'max_rhat'])

temp.bias$rhat <- as.numeric(rhats$rhat)
temp <- temp.bias %>% 
  reshape2::melt(id.vars = 'rhat') %>%
  transform(scenario = parse_number(gsub('S.', '', bias.files[f]))) %>%
  transform(value = ifelse(is.nan(value), 0, value))

rel.bias.20 <- rbind(rel.bias.20, temp)

} #rds file

rel.bias.20 <- rel.bias.20 %>%
  filter(rhat < 1.1) #%>%
  # transform(length = 20)

mean.rel.bias.20 <- data.frame(rel.bias.20) %>%
  group_by(variable, scenario) %>%
  dplyr::summarize(value = mean(value), .groups = 'keep') %>%
  merge(scenarios, by = 'scenario', all = F) %>%
  group_by(variable, detection, N, survey_freq) %>%
  dplyr::summarize(value = mean(value)*100, .groups = 'keep') %>%
  transform(variable = factor(variable, levels = c('phiP', 'phi1', 'phi2', 'phiJ', 'phiA',
                            'p1', 'p2', 'pJ', 'pA'))) %>%
  transform(detection = factor(detection, levels = c('Low', 'Med', 'High'))) %>%
  transform(survey_frequency = 'Branding frequency & study length', 
            samp_size = 'Sample size') %>%
  transform(length = 20) %>%
  bind_rows(mean.rel.bias)

#figures
#heatmaps
plot_dat <- mean.rel.bias.20 %>% filter(grepl('phi', variable)) %>%
  transform(variable = factor(variable, levels = c('phiP', 'phi1', 'phi2', 'phiJ', 'phiA'), 
                              labels = phi_pars))

#need to add "length" metric" or filter and do dot plot below
bias_heat <- ggplot(plot_dat, 
       aes(x = factor(detection), y = variable, fill = value)) +
  geom_tile(color = 'grey50') +
  xlab('Detection') + ylab('Parameter') +
  facet_nested(N ~ survey_freq + length, drop = T, 
             labeller = labeller(survey_freq = freq_labels, length = occ_labels),
             scales = 'free_x') +
  scale_fill_gradient2(name = "Relative bias (%)",
                      mid = "#FFFFFF", low = "deepskyblue1", high = "red2", midpoint = 0) +
  theme_bw() +
  theme(legend.position = 'right',
             strip.background = element_rect(fill = "#EEEEEE", color = "#FFFFFF"),
        legend.text = element_text(size = 8, vjust = 0.5),
        legend.title = element_text(vjust = 0.8, size = 10)) +
    scale_y_discrete("", labels = parse(text = c(eval(expression(levels(plot_dat$variable))))))

#dotplots
dotplot_dat <- rel.bias.20 %>% transform(length = 20) %>% 
  bind_rows(rel.bias %>% transform(length = 10)) %>%
  merge(scenarios, by = 'scenario', all = F) %>%
  group_by(length, variable, detection, survey_freq, N) %>%
  # dplyr::summarize(value = mean(value),
  #                  lower = quantile(value, probs = 0.025),
  #                  upper = quantile(value, probs = 0.975),
    dplyr::summarize(value = mean(value)*100, .groups = 'keep') %>%
  transform(variable = factor(variable, levels = c('phiP', 'phi1', 'phi2', 'phiJ', 'phiA'),
                              labels = c('Pup', 'Age-1', 'Age-2', 'Juvenile', 'Adult'))) %>%
  transform(detection = factor(detection, levels = c('Low', 'Med', 'High'),
                               labels = c('Low', 'Medium', 'High'))) %>%
  transform(survey_freq = factor(survey_freq, levels = c(1,2,3),
                                 labels = c('Annual', 'Biennial', 'Triennial'))) %>%
  transform(survey_frequency = 'Branding frequency', samp_size = 'Sample size',
            det = 'Detection') %>%
  #filter out detection params
  filter(!is.na(variable))

bias_dot <- ggplot(dotplot_dat %>% filter(N %in% c(50, 100, 200) & length == 20),
       aes(x = factor(survey_freq), y = value, col = factor(variable), group = variable)) +
  geom_point() + geom_line() +
  geom_hline(aes(yintercept = 0), linetype = 'dotted') +
  xlab('Branding frequency') + ylab('Relative bias') +
  facet_nested(det + detection ~ samp_size + N, drop = T,
             labeller = labeller(survey_freq = freq_labels, length = occ_labels),
             scales = 'free_x') +

  theme_bw() +
  plot_theme(legend.position = 'top',
             plot.subtitle = element_text(size = 10, hjust = 0.5, vjust = 1)) +
  scale_color_manual(values = rainbow2, name = 'Survival rate')


```

```{r RMSE}

rmse.files <- list.files(path = here::here('results', 'simpleAge', 'sims'), 
                    pattern = 'rmse', all.files = FALSE, full.names = F)

rmse <- data.frame()
for (f in 1:length(rmse.files)) {
temp.rmse <- data.frame(readRDS(file = here::here('results', 'simpleAge', 'sims',
                                                  rmse.files[f])))
#use gsub on rmse files rather than use different list to make sure they match
rhats <- data.frame(rhat = readRDS(file = here::here('results', 'simpleAge', 'sims', 
                                                     gsub('rmse', 'median', rmse.files[f])))[,'max_rhat'])

temp.rmse$rhat <- as.numeric(rhats$rhat)
temp <- temp.rmse %>% 
  reshape2::melt(id.vars = 'rhat') %>%
  transform(scenario = parse_number(gsub('S.', '', rmse.files[f]))) %>%
  transform(value = ifelse(is.nan(value), 0, value))

rmse <- rbind(rmse, temp)

} #rds file

rmse <- rmse %>%
  filter(rhat < 1.1) #%>%
  # transform(length = 10)

mean.rmse <- data.frame(rmse) %>%
  group_by(variable, scenario) %>%
  dplyr::summarize(value = mean(value), .groups = 'keep') %>%
  merge(scenarios, by = 'scenario', all = F) %>%
  group_by(variable, detection, N, survey_freq) %>%
  dplyr::summarize(value = mean(value), .groups = 'keep') %>%
  transform(variable = factor(variable, levels = c('phiP', 'phi1', 'phi2', 'phiJ', 'phiA',
                            'p1', 'p2', 'pJ', 'pA'))) %>%
  transform(detection = factor(detection, levels = c('Low', 'Med', 'High'))) %>%
  transform(survey_frequency = 'Branding frequency & study length', samp_size = 'Sample size') %>%
  transform(length = 10) 


rmse.files <- list.files(path = here::here('results', 'simpleAge', 'sims', 'prePop'), 
                    pattern = 'rmse', all.files = FALSE, full.names = F)

rmse.20 <- data.frame()
for (f in 1:length(rmse.files)) {
temp.rmse <- data.frame(readRDS(file = here::here('results', 'simpleAge', 'sims', 'prePop',
                                                  rmse.files[f])))
#use gsub on rmse files rather than use different list to make sure they match
rhats <- data.frame(rhat = readRDS(file = here::here('results', 'simpleAge', 'sims', 'prePop',
                                                     gsub('rmse', 'median', rmse.files[f])))[,'max_rhat'])

temp.rmse$rhat <- as.numeric(rhats$rhat)
temp <- temp.rmse %>% 
  reshape2::melt(id.vars = 'rhat') %>%
  transform(scenario = parse_number(gsub('S.', '', rmse.files[f]))) %>%
  transform(value = ifelse(is.nan(value), 0, value))

rmse.20 <- rbind(rmse.20, temp)

} #rds file

rmse.20 <- rmse.20 %>%
  filter(rhat < 1.1) #%>%
  # transform(length = 20) 

mean.rmse.20 <- data.frame(rmse.20) %>%
  group_by(variable, scenario) %>%
  dplyr::summarize(value = mean(value), .groups = 'keep') %>%
  merge(scenarios, by = 'scenario', all = F) %>%
  group_by(variable, detection, N, survey_freq) %>%
  dplyr::summarize(value = mean(value), .groups = 'keep') %>%
  transform(variable = factor(variable, levels = c('phiP', 'phi1', 'phi2', 'phiJ', 'phiA',
                            'p1', 'p2', 'pJ', 'pA'))) %>%
  transform(detection = factor(detection, levels = c('Low', 'Med', 'High'))) %>%
  transform(survey_frequency = 'Branding frequency & study length', samp_size = 'Sample size') %>%
  transform(length = 20) %>%
  bind_rows(mean.rmse)


#plots
plot_dat <- mean.rmse.20 %>% filter(grepl('phi', variable)) %>%
  transform(variable = factor(variable, levels = c('phiP', 'phi1', 'phi2', 'phiJ', 'phiA'), 
                              labels = phi_pars))

rmse_heat <- ggplot(plot_dat, aes(x = factor(detection), y = variable, fill = value)) +
  geom_tile(color = 'grey50') +
  xlab('Detection') +
  ylab('Parameter') +
  facet_nested(samp_size + N ~ survey_frequency + survey_freq + length, drop = T, 
             labeller = labeller(survey_freq = freq_labels, length = occ_labels),
             scales = 'free_x') +
 scale_fill_gradient2(name = "RMSE",
                      mid = "#FFFFFF", low = "#012345", high = "#012345", midpoint = 0) +
  theme_bw() +
  theme(legend.position = 'right',
             strip.background = element_rect(fill = "#EEEEEE", color = "#FFFFFF"),
        legend.text = element_text(size = 8, vjust = 0.5),
        legend.title = element_text(vjust = 0.8, size = 10)) +
    scale_y_discrete("", labels = parse(text = c(eval(expression(levels(plot_dat$variable))))))

##dot plots
dotplot_dat <- rmse.20 %>% transform(length = 20) %>% 
  bind_rows(rmse %>% transform(length = 10)) %>%
  merge(scenarios, by = 'scenario', all = F) %>%
  group_by(length, variable, detection, survey_freq, N) %>%
  # dplyr::summarize(value = mean(value),
  #                  lower = quantile(value, probs = 0.025),
  #                  upper = quantile(value, probs = 0.975),
    dplyr::summarize(value = mean(value), .groups = 'keep') %>%
  transform(variable = factor(variable, levels = c('phiP', 'phi1', 'phi2', 'phiJ', 'phiA'),
                              labels = c('Pup', 'Age-1', 'Age-2', 'Juvenile', 'Adult'))) %>%
  transform(detection = factor(detection, levels = c('Low', 'Med', 'High'),
                               labels = c('Low', 'Medium', 'High'))) %>%
  transform(survey_freq = factor(survey_freq, levels = c(1,2,3),
                                 labels = c('Annual', 'Biennial', 'Triennial'))) %>%
  transform(survey_frequency = 'Branding frequency', samp_size = 'Sample size',
            det = 'Detection') %>%
  #filter out detection params
  filter(!is.na(variable))

rmse_dot <- ggplot(dotplot_dat %>% filter(N %in% c(50, 100, 200) & length == 20),
       aes(x = factor(survey_freq), y = value, col = factor(variable), group = variable)) +
  geom_point() + geom_line() +
  geom_hline(aes(yintercept = 0), linetype = 'dotted') +
  xlab('Branding frequency') + ylab('RMSE') +
  facet_nested(detection ~ samp_size + N, drop = T,
             labeller = labeller(survey_freq = freq_labels, length = occ_labels),
             scales = 'free_x') +

  theme_bw() +
  plot_theme(legend.position = 'top',
             plot.subtitle = element_text(size = 10, hjust = 0.5, vjust = 1)) +
  scale_color_manual(values = rainbow2, name = 'Rate')

```


### Introduction (outline)  

Important concepts/paragraphs

- Accurate/unbiased estimates are important for wildlife conservation and management
-- Status assessments (abundance), effects of changing environmental conditions (demography)
-- Have to achieve a certain degree of precision to be effective in meeting conservation and management goals (Reynolds et al)
-- These estimates can be expensive, invasive, disruptive, logistically challenging
-- Important to "optimize" trade-offs for managers, wildlife, and costs 

- Optimizing can be logistically challenging 
-- Costs can be difficult to quantify
-- Budgets can be difficult to predict/changing
-- Can change as project needs are distilled or population status changes over long-term
-- Some levers can be pulled, some can't
-- Metrics/goals might not be well defined
-- Usually not done before a sampling design, but just to ensure no concerning bias at given study features (Yackulic et al. 2018)
--- Examples

- Optimizing can be subjective  
-- Political/controversial/different stakeholders
-- Risk tolerance is usually difficult to elicit or poorly defined  
--- Examples lead into SSL

- Even if it is hard, it is worth doing
-- Benefits for SSL
-- Applicability for other populations
-- Conclusion sentence

### Methods  
-- Study system: briefly explain layout of different rookeries/regions/trends?
 
-- Model description and equations  
 
-- Simulation scenarios
 
-- Performance metrics tracked  
 
### Results  
 
### Discussion  
 
- Aims/results/methods review
-- Review most compelling results
-- Summarize benefits/applications/importance
 
- Review other studies that have done this?
-- Example 1
-- Example 2
 
- What it means for SSL  
-- Study design must also consider "risk tolerance" and management goals

-- What are those for SSL? They vary across space (and time as management needs change?)
 
-- Goals have to be re-evaluated in iterative process

- Recommendations for mark-recapture designs generally
-- Example 1
 
- Conclusion
-- Benefits and challenges of study
-- Important for future of dwindling funding



### Literature review

Pollock et al. 1990 
- Statistical inference for capture-recapture experiments. Wildlife Monographs 107:3–97.
- recommends CV < 0.2? (Kristensen & Kovach 2018)

Campbell et al. 2002
- an assessment of monitoring efforts in endangered species recovery plans; a review of 181 species
- were proposed tasks implemented?
- Trend monitoring/abundance most commonly proposed recovery metric, largely because oftentimes recovery has to be initiated in the face of incomplete information and other barriers
- Not always easy to prioritize implementing "monitoring" if other efforts (threat mitigation, captive breeding, habitat restoration) might have more direct effects on abundance
- Species-specific monitoring plans are important 
- Proposed/implemented monitoring generally didn't address threats

Naidoo et al. 2006
- Incorporating costs into conservation planning
- By incorporating spatial distribution of benefits (not all areas provide equal benefits), you get a more efficient and effective outcome per dollar if account for (spatial heterogeneity of) costs
- Types of different costs - numerous
- Barriers: (1) biologists "invented" field of conservation and aren't trained as economists, and (spatially explicit) cost data isn't available, but number of studies are increasing; (2) many conservation plans are just difficult/infeasible to implement; (3) ecologists reluctant to include factors other than biological (economic) in conservation prioritization
- Nice quote: "Balancing research on biodiversity features (i.e. the benefits side) with a greatly strengthened understanding of economic (and indeed other) aspects of the costs side will lead to novel and creative ways to obtain environmental benefits in the most efficient manner possible."

Field et al. 2005
Field, S.A., Tyre, A.J., Possingham, H.P., 2005. Optimizing allocation of monitoring effort under economic and observational constraints. Journal of Wildlife Management 69, 473–482.

Nichols & Williams. 2006
- Monitoring for conservation
- Framing comparing to Platt's original criticism of collecting scientific data without clear hypothesis-testing application/goals; likens that to targeted monitoring vs. "omnimbus surveillance monitoring"; define targeted as being "integrated into conservation practice" alongside monitoring design and implementation informed by hypotheses about ecological responses to management actions
- Surveillance often implemented to "learn more about a system", which can be helpful, but doesn't get at effective and efficiency criteria
- Conservation "rooted in decision theory"; the need for making state-dependent decisions (includes confidence about model, quantifying uncertainty, etc)
- Informed decision processes and adaptive management: objectives, management actions, model of response to management action, measure of confidence, and monitoring program that provides estimates of system state
- Understandable that "omnibus" monitoring is so prevalent - could arguably be most versatile as questions arise and cover most species? not arguing fully against it, just that it might not be the most efficient
- Surveillance often triggered by identifying a decline; better if possible to understand best remedy for recovery rather than reason for decline
- Particularly important in a time with limited conservation funding and such widespread conservation need 
- Very nice paper

Field et al. 2007
Field, S.A., O’Connor, P.J., Tyre, A.J., Possingham, H.P., 2007. Making monitoring meaningful. Austral Ecology 32, 485–491. 
- Estimates need to be accurate/precise to be useful/answer the right questions
- Citation for not commonly done?

Taylor et al. 2007
Lessons from monitoring trends in abundance of marine mammals. Marine Mammal Science 23, 157175.
- Failure to plan wastes valuable resources

Lyons et al. 2008
Lyons, J.E., Runge, M.C., Laskowski, H.P., Kendall, W.L., 2008. Monitoring in the context of structured decision-making and adaptive management. Journal of Wildlife Management 72, 1683–1692.
- Information should be useful, i.e. affect decision-making; wastes resources if fail to improve the process and miss documenting things you intended to

Reynolds et al. 2011
- Nice intro framing
- Identifying effective and efficient survey designs for monitoring
- Monitoring program only "effective" if it produces info with sufficient accuracy/precision to answer scientific questions and influence decisions
- Generally, speed at detecting a change is important, but resources constrain, so need efficient: quickest detection for lowest cost (*Sims et al. 2008*)
- Effectiveness and efficiency determined by survey design; sampling effort (per survey) and survey frequency main components of this effective-efficient trade-off spectrum
- Despite importance, not often done (Legg & Nagy 2006, Lindenmayer & Likens 2009)
- Risks: resources that could have gone elsewhere are wasted, and program's efficacy/ineffectiveness may be unresolved due to natural variability in the state process, and lose opportunity to have well-defined management objectives, loss of long-term institutional support/funding; can create a "falst sense of management performance", doesn't promote true understanding of costs of (helpful) information
- Planning becomes more important and more complex with increasingly complex observation processes (sources of bias, logistical constraints, multiple sources of variability)
- Applied to brown bears in Alaska
- Management objectives, actions, and information needs

Sollmann et al. 2012
- Black bear data and simulation study; not much discussion about simulation

Torrubia et al. 2014
- Getting most connectivity per conservation dollar; Washington ground squirrel (Urocitellus washingtoni)
- Found that incorporating spatially explicit cost information improved efficiency (reduced overall costs and increase % land restored) of conservation actions
- Very land/terrestrial/connectivity-focused

Sun et al. 2014
- Trap configuration in spatial MR; black bear (ursus americanus)
- Different configurations (spacing and number of traps) work, but highly dependent on home range size
- Varied detection and spatial scaling parameters (how much animal travels relative to home range?)

Peel et al. 2014
- Mark-recapture simulations for acoustic abundance estimates of whales (Antactic blue)
- Lower encounter rates for rare species (line transects and mark-resight)
- Aimed to assist in planning mark-recap surveys for whales (pre-planning side of things)
- Densely detailed information about encounter rates, vocalization/movement assumptions; not a clear explanation of the survey features that they were examining across.... mainly showed that the acoustic assistance improves estimation over mark-recapture alone

Wilton et al. 2014
- Trap precision configuration influences black bear density/abundance estimation
- Mainly tested bias under proposed study designs, but found that situations with low density and non-uniform distribution required more thoughtful trade-offs about snare spacing, coverage, and sample sizes to achieve precision

Johnston et al. 2015
- Abundance models (compared to occupancy) improve spatial and temporal prioritization of conservation resources; not a simulation study, probably don't need to cite

Peel et al. 2015
- Same study/issue as above, better citation - more focused on study design
- Estimates of abundance have to be appropriately precise to be useful (Reynolds et al 2011)
- Marine mammal surveys (particularly for dispersed and remote areas) can be challenges and expensive (Williams and Thomas 2009), making planning really important to make sure resources aren't wasted
- CV across true mortality, sex ratio, model, and growth rates, and survey features (number of vessels, survey length, distribution of effort)
- Pre-survey planning simulations are important to ensure that investment meets desired precision and is able to answer the questions being asked

Yackulic et al. 2015
- Modeled species interactions and simulated different underlying state processes to examine when bias might be an issue (citation for examining specific situation rather than using to design study beforehand)

Burton et al. 2015
- Linking camera trapping to ecological processes; review study examining number of papers not accounting appropriately for all the potential biases given ecological processes, but good citation for proliferation/adoption of camera trapping technology

Lyons et al. 2015
- Superpopulation size, proportion marked, and number of animals sampled (stopover duration for red knot sandpipers; Calidris canutus roselaari)
- Relative bias less affected by number of birds per scan sample or superpopulation size
- Accuracy (as measured by RMSE) improved with more scans per day and with larger scan sample sizes; same with CV

Evans et al. 2016. 
- Species recovery in the United States: increasing the effectiveness of the ESA
- Insufficient funding for species recovery, and not proportional across taxa (mostly for fish; salmon/sturgeon)
- Funding shortages bring trade-offs front and center; allocations more often driven by political/social factors
- Mitigations necessary for climate change will likely further constrain recovery resources
- Limited resources often preclude adaptive management feedback applications
- Hard to know how often/if the agencies follow their own prioritization scheme for allocating recovery funds

Galvez et al. 2016
- Incorporating costs into camera trap surveys
- "Paucity of research examining how to allocate survey effort to optimize estimation that takes into account operational costs"
- Incorporate costs per sampling unit, then look across survey lengths and number of cameras needed to achieve precision targets given detection/occupancy scenarios
- Generally more cost-efficient to do multiple camera traps over fewer occasions because travel/salaries are big part of costs, particularly for elusive species

Gonzalez et al. 2016
- UAVs and AI monitoring for the future

Lanier et al. 2016
- Effects of survey design on demography, simulation for frogs, open robust design
- Synchronous vs. asynchronous breeding, logistical access (whether breeding grounds can be surveyed)
- Fairly large range of bias in survival across scenarios

Little et al. 2016
- trade-offs for cost-constrained fisheries management
- lots of lists of all the different "risks" that incur more costs (e.g., the need to implement a control rule for additional cost)
- risk (of overfishing)-catch-cost frontier.... easier to derive benefits for fisheries compared to wildlife monitoring

Sackett & Catalano 2017
- Red snapper (Lutjanus campechanus) in GoM; recreational fishing mortality; sample size (number tagged), fishing effort, fish density, tag loss - lots of facets
- Results included "most cost effective study design"
- High vs. variable reward fish tags
- Precision improved with increasing sample sizes, but so does costs
- Lots of details testing specific study designs
- Have to consider incentive vs. reporting rate

Armsworth et al. 2017
- Argue that factoring in economic costs doesn't *always* improve efficiency - sometimes can exacerbate disagreements about priorities.... still have to be very specific about protection efforts and how to quantify benefits
- Disagreements are worst when benefits data are more variable
- Benefits of accounting for cost are more variable when cost data are more variable
- Simulation approach

Del Vecchio et al. 2018
- Trade-offs between effort and data quality - habitat monitoring
- Plot sampling schemes; meh

Kristensen & Kovach 2018
- SECR and new england cottontail (Sylvilagus transitionalis)
- Studies are less common for smaller individuals; conclude optimal study designs will vary among organisms (habitat use and spatial extent), so pilot planning is important
- *Campbell et al. 2002* abundance is the most common quantitative metric in recovery plans and delisting criteria, key parameter underlying extinction risk
- *Need good abundance estimates to measure response to management actions (Nichols & Williams 06; Johnston et al. 2015)*
- CV and relative bias; number of survey occasions, pellet collection spacing, true density, transect spacing all important predictors; number of survey occasions less impactful?
- "When detectability was low (g0 = 0.1), both additional detectors and more frequent sampling were required to achieve the desired levels of precision than with higher detectability (g0 = 0.2)"
- *Pollock et al 1990* recommends CV < 0.2? 
- Also found that true density had large effect on performance, had to compensate for low density by upping observation metrics
- Nice examples of when management goals affect needed precision: tracking fate of reintroduction (maybe need to ensure less bias and higher precision) compared to responses to habitat management (number alive with low precision acceptable?)

Durden et al. 2021
- Bottlenose dolphine (Tursiops truncatus truncatus) in Indian River Lagoon
- Robust design and survey optimization simulation; limited description of results - mainly validating their design - checking precision; but also looking for ways to reduce survey effort without losing precision, and suggest importance of long-term planning

Freeman et al. 2021
- SECR on small mammals; number of trapping occasions, trap number - both metrics improved estimation; salt marsh harvest mouse (Reithrodontomys raviventris)

White et al. 2022
- Decision support framework for assessing costs and benefits of conservation actions
- Need for incorporating costs, but difficult to estimate and poorly/infrequently reported and therefore rarely studied
- Mainly a how-to


### Literature cited






