---
title: "SSL Future Proof Project Update - August 1 2022"
output: html_document
---

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = FALSE, eval = T, message = F, warning = F)

library(here) #v1.0.1
library(dplyr) #v1.0.5
library(nimble) #v0.12.1
library(coda)
library(ggh4x)
library(readr)
library(knitr)
library(Hmisc)

source(here::here('scripts', 'PlotTheme.R'))

```

### Demographic and detection rate estimates  

Survival rates were estimated for ages P, 1:9, and age 10+, with fixed effects estimating the effect of pup mass and the differences between sex and natal region on survival over 19 years of data. The sample sizes for natal region were imbalanced, with more samples from Pyramid Rock (*n* = 1395) than South Seal Rock (*n* = 290). Fixed effects of sex were estimated for each age group while those of natal region were estimated for pups, yearlings, shared for juveniles (ages 3-6) and adults (ages 7-10+). Temporal variance was estimated using a random effect of year with a penalized complexity prior, with the random effect estimated for pups, yearlings, and pooled for juveniles (ages 3-6) and adults (ages 7-10+) jointly for both sex and natal region (i.e., trends were not estimated for each sex nor region but for age groups in the eDPS as a whole). 

Detection rates were estimated for each age as above with temporal variance constructed in the same fashion, but detection was assumed to be constant across sexes, natal regions, and resight location. 

#### Survival rates  

```{r results - phi}

out <- readRDS(here::here('results', 'out_ME_fixed_sd_switch.RDS'))

all_pars <- colnames(out$samples$chain1)
noZ <- all_pars[which(!grepl('z', all_pars))]

outmat <- as.matrix(out$samples)
posts <- outmat[,noZ]

post_sum <- data.frame(
  med = apply(posts, 2, function(x) quantile(x, probs = 0.5, na.rm = T, names = F)),
  lower = apply(posts, 2, function(x) quantile(x, probs = 0.025, na.rm = T, names = F)),
  upper = apply(posts, 2, function(x) quantile(x, probs = 0.975, na.rm = T, names = F)))
post_sum$variable <- row.names(post_sum)

#mean rates
sex.pars <- all_pars[which(grepl('b.sex', all_pars))]
mu.pars <- all_pars[which(grepl('mu.', all_pars))]

pars <- c(sex.pars, mu.pars)

mu.vals <- post_sum %>%
  filter(variable %in% mu.pars) %>%
  transform(age = gsub('mu.', '', variable)) 

b.sex <- post_sum %>% filter(grepl('b.sex', variable) & med != 0) %>%
  transform(age = ifelse(grepl('sexP', variable), 'P', 
                         ifelse(grepl('sex1', variable), '1',
                                ifelse(grepl('sex2', variable), '2', 
                                       ifelse(grepl('sex3', variable), '3', 
                                              ifelse(grepl('sex4', variable), '4',
                                                     ifelse(grepl('sex5', variable), '5',
                                                            ifelse(grepl('sex6', variable), '6', 'A'))))))))

#male back transformed values
mean.phi.vals <- mu.vals %>%
  merge(b.sex, by = 'age', suffixes = c('', '.sex'), all = T) %>%
  #fill b.sex for older age intercepts
  transform(med.sex = ifelse(is.na(med.sex), b.sex[b.sex$age == 'A', 'med'], med.sex),
            lower.sex = ifelse(is.na(lower.sex), b.sex[b.sex$age == 'A', 'lower'], lower.sex),
            upper.sex = ifelse(is.na(upper.sex), b.sex[b.sex$age == 'A', 'upper'], upper.sex)) %>%
  transform(med = 1/(1+exp(-(med + med.sex))),
            lower = 1/(1+exp(-(lower + lower.sex))),
            upper = 1/(1+exp(-(upper + upper.sex)))) %>%
  transform(sex = 'Male') %>%
  dplyr::select(c(med, lower, upper, variable, age, sex)) %>%
  bind_rows(
    post_sum %>%
  filter(grepl('int.phi', variable)) %>%
  transform(age = gsub('int.phi', '', variable), sex = 'Female')) %>%
  transform(age = factor(age, levels = c('P', '1', '2', '3', '4', '5', '6', '7', '8', '9', 'A')))

ggplot(mean.phi.vals, aes(age, med, col = sex, group = sex)) +
  geom_point(position = position_dodge(0.5)) +
  geom_errorbar(aes(ymin = lower, ymax = upper), width = 0.5, position = position_dodge(0.5)) +
  xlab('') + ylab(expression(paste('Mean survival probability ',  '(', phi, ')'))) +
  ggtitle('') +
  plot_theme(legend.position = 'top', panel.border = element_rect(fill = NA),
             plot.title = element_text(hjust = 0.5)) +
  guides(color = guide_legend("", nrow = 1, byrow = T)) +
  scale_color_manual(values = rainbow2[c(2,5)], name = '') 

#temporal variance
phi.vals <- post_sum %>%
  filter(grepl('.prob', variable)) %>%
  transform(year = as.numeric(as.character(gsub('.+\\[([0-9]+), ([0-9]+)\\].*$', '\\2', 
                                                variable)))) %>%
  transform(region = gsub('.+\\[([0-9]+), ([0-9]+)\\].*$', '\\1', variable)) %>%
  transform(variable = sub("\\[.*", "", variable)) %>%
  transform(age = gsub('.prob', '', variable)) %>%
  transform(age = gsub('phi', '', age)) %>%
  transform(age_cat = factor(age, levels = c('P', '1', '2', '3', '4', '5', '6', '7', '8', '9', 'A',
                                         'PM', '1M', '2M', '3M', '4M', '5M', '6M', '7M', '8M', 
                                         '9M', 'AM'))) %>%
  transform(sex = ifelse(grepl('M', age), 'Male', 'Female')) %>%
  transform(age = gsub('M', '', age)) %>%
  transform(age = factor(age, levels = c('P', '1', '2', '3', '4', '5', '6', '7', '8', '9', 'A'))) %>%
  merge(mean.phi.vals, by = c('age', 'sex'), suffixes = c('', '.mean'))  %>%
  transform(region = factor(region, levels = c(1,2), 
                            labels = c('Pyramid rock', 'South seal rock')))

#temporal variance with mean rates; intercept represents female, so can't show males
# ggplot(phi.vals %>% filter(sex == 'Female' & region == 'Pyramid rock'),
#        aes(year, med, color = factor(region), group = factor(region))) +
#   geom_errorbar(aes(x = year, ymin=lower, ymax=upper), position = position_dodge(0.5),
#                 width = 0.5, show.legend = F) +
#   geom_point(size = 0.8, position = position_dodge(0.5)) +
#   geom_line(size = 0.7, position = position_dodge(0.5)) +
#   geom_line(aes(year, med.mean), col = 'grey20', linetype = 'dotted') +
#   xlab('') + ylab(expression(paste('Predicted survival probability ',  '(', phi, ')'))) +
#   ggtitle('') +
#     facet_grid(. ~ age) +
#   # facet_grid(age ~ region) +
#   plot_theme(legend.position = 'top', panel.border = element_rect(fill = NA),
#              plot.title = element_text(hjust = 0.5)) +
#   guides(color = guide_legend("", nrow = 1, byrow = T)) +
#   scale_color_manual(values = rainbow2[c(2,3,5,6,7,8)], name = '') +
#   scale_x_continuous(breaks = c(seq(1, 19, by = 2)), labels = c(seq(2001, 2019, by = 2)))

#female survival across both regions
ggplot(phi.vals %>% filter(sex == 'Female'),
       aes(year, med, color = factor(region), group = factor(region))) +
  geom_errorbar(aes(x = year, ymin=lower, ymax=upper), position = position_dodge(0.5), 
                width = 0.5, show.legend = F) +
  geom_point(size = 0.8, position = position_dodge(0.5)) + 
  geom_line(size = 0.7, position = position_dodge(0.5)) +
  xlab('') + ylab(expression(paste('Predicted survival probability ',  '(', phi, ')'))) +
  ggtitle('') +
    facet_wrap(. ~ age) +
  # facet_grid(age ~ region) +
  plot_theme(legend.position = 'top', panel.border = element_rect(fill = NA),
             plot.title = element_text(hjust = 0.5)) +
  guides(color = guide_legend("", nrow = 1, byrow = T)) +
  scale_color_manual(values = rainbow2[c(2,3,5,6,7,8)], name = '') +
  scale_x_continuous(breaks = c(seq(1, 19, by = 2)), labels = c(seq(2001, 2019, by = 2)))

#just young individuals
# ggplot(phi.vals %>% filter(sex == 'Female' & age %in% c('P', '1', '2')),
#        aes(year, med, color = factor(region), group = factor(region))) +
#   geom_errorbar(aes(x = year, ymin=lower, ymax=upper), position = position_dodge(0.5), 
#                 width = 0.5, show.legend = F) +
#   geom_point(size = 0.8, position = position_dodge(0.5)) + 
#   geom_line(size = 0.7, position = position_dodge(0.5)) +
#   xlab('') + ylab(expression(paste('Predicted survival probability ',  '(', phi, ')'))) +
#   ggtitle('') +
#     facet_grid(. ~ age) +
#   # facet_grid(age ~ region) +
#   plot_theme(legend.position = 'top', panel.border = element_rect(fill = NA),
#              plot.title = element_text(hjust = 0.5)) +
#   guides(color = guide_legend("", nrow = 1, byrow = T)) +
#   scale_color_manual(values = rainbow2[c(2,3,5,6,7,8)], name = '') +
#   scale_x_continuous(breaks = c(seq(1, 19, by = 2)), labels = c(seq(2001, 2019, by = 2)))


```

#### Detection rates 
```{r results - p}

#mean detection probabilities
p.vals <- post_sum %>%
  filter(!grepl('phi', variable) & grepl('int.p', variable)) %>%
  transform(age = gsub('int.p', '', variable)) %>%
  transform(age = factor(age, levels = c('1', '2', '3', '4', '5', '6', '7', '8', '9', 'A')))

ggplot(p.vals, aes(age, med)) +
  geom_point() +
  geom_errorbar(aes(ymin = lower, ymax = upper)) +
  xlab('') + ylab('Detection probability') +
  ggtitle('') +
  plot_theme(legend.position = 'top', panel.border = element_rect(fill = NA),
             plot.title = element_text(hjust = 0.5)) +
  guides(color = guide_legend("", nrow = 1, byrow = T)) +
  scale_color_manual(values = rainbow2[c(2,3,5,6,7,8)], name = '') 

#temporal variance
p.vals <- post_sum %>%
  filter(grepl('.prob', variable)) %>%
  filter(!grepl('phi', variable)) %>%
  transform(year = as.numeric(as.character(gsub('.+\\[([0-9]+)\\].*$', '\\1', 
                                                variable)))) %>%
  transform(variable = sub("\\[.*", "", variable)) %>%
  transform(age = gsub('.prob', '', variable)) %>%
  transform(age = gsub('p', '', age)) %>%
  transform(age = factor(age, levels = c('1', '2', '3', '4', '5', '6', '7', '8', '9', 'A')))

ggplot(p.vals, aes(year, med)) +
  geom_errorbar(aes(x = year, ymin=lower, ymax=upper), position = position_dodge(0.5), 
                width = 0.5, show.legend = F) +
  geom_point(size = 0.8, position = position_dodge(0.5)) + 
  geom_line(size = 0.7, position = position_dodge(0.5)) +
  xlab('') + ylab(expression(paste('Predicted detection probability '))) +
  ggtitle('') +
    facet_wrap(. ~ age) +
  # facet_grid(age ~ region) +
  plot_theme(legend.position = 'top', panel.border = element_rect(fill = NA),
             plot.title = element_text(hjust = 0.5)) +
  guides(color = guide_legend("", nrow = 1, byrow = T)) +
  scale_color_manual(values = rainbow2[c(2,3,5,6,7,8)], name = '') +
  scale_x_continuous(breaks = c(seq(1, 19, by = 2)), labels = c(seq(2001, 2019, by = 2)))

```

### Simulation analysis

The above model was fit using simulated data across 12 scenarios with varying sample sizes (N = 50, 200), branding frequency (annual, biennial), and detection rates (low, medium, high; low and high = +/- 20% of medium). The only difference between the model in the simulation versus that fit with real data is grouped mean detection rates (pups, yearlings, juveniles, adults). No temporal variance estimated (yet). Model performance was (so far) examined using relative bias and RMSE. All parameters in all simulations (*n* = 50) converged (Rhat < 1.05). 

```{r convergence summary}

sc <- c(1:12)

scenarios <- data.frame(scenario = c(sc),
                        survey_freq = c(rep(1,6), rep(2,6)),
                        detection = c(rep(c('Low', 'Low', 'Med', 'Med', 'High', 'High'), 2)),
                        N = c(rep(c(50,200), 6)))
                        
med.files <- list.files(path = here::here('results', 'sims'), 
                    pattern = 'median', all.files = FALSE, full.names = F)


converged <- data.frame()
for (f in 1:length(med.files)) {

rhats <- data.frame(rhat = readRDS(file = here::here('results', 'sims', 
                                                     med.files[f]))[,'max_rhat']) %>%
  transform(scenario = parse_number(gsub('S.', '', med.files[f]))) %>%
  transform(rhat = as.numeric(rhat))

converged <- rbind(converged, rhats)

}

tot_runs <- converged %>%
  filter(!is.na(rhat)) %>%
  group_by(scenario) %>%
  dplyr::summarize(tot = n())

prop_conv <- converged %>%
  filter(rhat<1.10) %>%
  group_by(scenario) %>%
  dplyr::summarize(conv = n()) %>%
  merge(tot_runs, by = 'scenario', all = T) %>%
  transform(prop_conv = round(conv/tot,1)) %>%
  dplyr::select(scenario, tot, prop_conv)

scenarios_tab <- scenarios %>%
  merge(prop_conv, by = 'scenario') %>%
  transform(survey_freq = factor(survey_freq, levels = c(1,2),
                                 labels = c('Annual', 'Biennial')))

kable(scenarios_tab)


```

#### Relative bias 

Plots showing relative bias for survival and detection across 12 simulation scenarios. 

```{r results bias}

bias.files <- list.files(path = here::here('results', 'sims'), 
                    pattern = 'bias', all.files = FALSE, full.names = F)

rel.bias <- data.frame()
for (f in 1:length(bias.files)) {
temp.bias <- data.frame(readRDS(file = here::here('results', 'sims',
                                                  bias.files[f])))
#use gsub on bias files rather than use different list to make sure they match
rhats <- data.frame(rhat = readRDS(file = here::here('results', 'sims',
                                                     gsub('bias', 'median', bias.files[f])))[,'max_rhat'])

temp.bias$rhat <- as.numeric(rhats$rhat)
temp <- temp.bias %>% 
  reshape2::melt(id.vars = 'rhat') %>%
  transform(scenario = parse_number(gsub('S.', '', bias.files[f]))) %>%
  transform(value = ifelse(is.nan(value), 0, value))

rel.bias <- rbind(rel.bias, temp)

} #rds file

rel.bias <- rel.bias %>%
  filter(rhat < 1.1) 

mean.rel.bias <- data.frame(rel.bias) %>%
  group_by(variable, scenario) %>%
  dplyr::summarize(value = mean(value), .groups = 'keep') %>%
  merge(scenarios, by = 'scenario', all = F) %>%
  group_by(variable, detection, N, survey_freq) %>%
  dplyr::summarize(value = mean(value)*100, .groups = 'keep') %>%
  transform(variable = factor(variable, levels = c('phiP', 'phi1', 'phi2', 'phi3', 'phi4',
                            'phi5', 'phi6', 'phi7', 'phi8', 'phi9', 'phiA',
                            'p1', 'p2', 'pJ', 'pA'))) %>%
  transform(detection = factor(detection, levels = c('Low', 'Med', 'High'))) %>%
  transform(survey_frequency = 'Survey frequency', samp_size = 'Sample size')

phi_pars <- c('phi[P]','phi[1]', 'phi[2]', 'phi[3]', 'phi[4]','phi[5]',
              'phi[6]','phi[7]', 'phi[8]', 'phi[9]', 'phi[10^plus]')
p_pars <- c('p[1]', 'p[2]', 'p[J]', 'p[A]') 


freq_labels <- c('1' = 'Annual', '2' = 'Biennial')

#phi
plot_dat <- mean.rel.bias %>% filter(grepl('phi', variable)) %>%
  transform(variable_lab = factor(variable, levels = c('phiP', 'phi1', 'phi2', 'phi3', 'phi4',
                                                   'phi5', 'phi6', 'phi7', 'phi8', 'phi9', 'phiA'), 
                              labels = phi_pars))

ggplot(plot_dat, aes(x = factor(detection), y = variable, fill = value)) +
  geom_tile(color = 'grey50') +
  xlab('Detection probability') +
  ylab('Parameter') +
  facet_nested(samp_size + N ~ survey_frequency + survey_freq, drop = T, 
             labeller = labeller(survey_freq = freq_labels),
             scales = 'free_x') +
  scale_fill_gradient2(name = "Relative bias (%)",
                      mid = "#FFFFFF", low = "deepskyblue1", high = "red2", midpoint = 0) +
  theme_bw() +
  theme(legend.position = 'right',
             strip.background = element_rect(fill = "#EEEEEE", color = "#FFFFFF"),
        legend.text = element_text(size = 8, vjust = 0.5),
        legend.title = element_text(vjust = 0.8, size = 10)) +
    scale_y_discrete("", labels = parse(text = c(eval(expression(levels(plot_dat$variable_lab))))))

#without older ages
ggplot(plot_dat %>% filter(variable %nin% c('phiA', 'phi9', 'phi8', 'phi7')), aes(x = factor(detection), y = variable, fill = value)) +
  geom_tile(color = 'grey50') +
  xlab('Detection probability') +
  ylab('Parameter') +
  facet_nested(samp_size + N ~ survey_frequency + survey_freq, drop = T,
             labeller = labeller(survey_freq = freq_labels),
             scales = 'free_x') +
  # facet_grid(N ~ survey_freq, drop = T,
  #            # labeller = labeller(survey_freq = freq_labels),
  #            scales = 'free_x') +
  scale_fill_gradient2(name = "Relative bias (%)",
                      mid = "#FFFFFF", low = "deepskyblue1", high = "red2", midpoint = 0) +
  theme_bw() +
  theme(legend.position = 'right',
             strip.background = element_rect(fill = "#EEEEEE", color = "#FFFFFF"),
        legend.text = element_text(size = 8, vjust = 0.5),
        legend.title = element_text(vjust = 0.8, size = 10)) +
    scale_y_discrete("", labels = parse(text = c(eval(expression(levels(plot_dat$variable_lab))))))


#just younger ages
ggplot(plot_dat %>% filter(variable %in% c('phi[P]', 'phi[1]', 'phi[2]', 'phi[3]', 'phi[4]',
                                           'phi[5]')), 
       aes(x = factor(detection), y = variable, fill = value)) +
  geom_tile(color = 'grey50') +
  xlab('Detection') +
  ylab('Parameter') +
  facet_nested(samp_size + N ~ survey_frequency + survey_freq, drop = T, 
             labeller = labeller(survey_freq = freq_labels),
             scales = 'free_x') +
  scale_fill_gradient2(name = "Relative bias (%)",
                      mid = "#FFFFFF", low = "deepskyblue1", high = "red2", midpoint = 0) +
  theme_bw() +
  theme(legend.position = 'right',
             strip.background = element_rect(fill = "#EEEEEE", color = "#FFFFFF"),
        legend.text = element_text(size = 8, vjust = 0.5),
        legend.title = element_text(vjust = 0.8, size = 10),
        axis.title.x = element_blank()) +
    scale_y_discrete("", labels = parse(text = c(eval(expression(levels(plot_dat$variable))))))

#p
plot_dat <- mean.rel.bias %>% filter(variable %in% c('p1', 'p2', 'pJ', 'pA')) %>%
  transform(variable = factor(variable, 
                              levels = c('p1', 'p2', 'pJ', 'pA'), 
                              labels = p_pars))

ggplot(plot_dat, aes(x = factor(detection), y = variable, fill = value)) +
  geom_tile(color = 'grey50') +
  xlab('Detection probability') +
  ylab('Parameter') +
  facet_nested(samp_size + N ~ survey_frequency + survey_freq, drop = T, 
             labeller = labeller(survey_freq = freq_labels),
             scales = 'free_x') +
  scale_fill_gradient2(name = "Relative bias (%)",
                      mid = "#FFFFFF", low = "deepskyblue1", high = "red2", midpoint = 0) +
  theme_bw() +
  theme(legend.position = 'right',
             strip.background = element_rect(fill = "#EEEEEE", color = "#FFFFFF"),
        legend.text = element_text(size = 8, vjust = 0.5),
        legend.title = element_text(vjust = 0.8, size = 10)) +
    scale_y_discrete("", labels = parse(text = c(eval(expression(levels(plot_dat$variable))))))


```

#### Root mean square error (RMSE)  

Plots showing RMSE for survival and detection across 12 simulation scenarios. 

```{r results rmse}

rmse.files <- list.files(path = here::here('results', 'sims'), 
                    pattern = 'rmse', all.files = FALSE, full.names = F)

rmse <- data.frame()
for (f in 1:length(rmse.files)) {
temp.rmse <- data.frame(readRDS(file = here::here('results', 'sims',
                                                  rmse.files[f])))
#use gsub on rmse files rather than use different list to make sure they match
rhats <- data.frame(rhat = readRDS(file = here::here('results', 'sims',
                                                     gsub('rmse', 'median', rmse.files[f])))[,'max_rhat'])

temp.rmse$rhat <- as.numeric(rhats$rhat)
temp <- temp.rmse %>% 
  reshape2::melt(id.vars = 'rhat') %>%
  transform(scenario = parse_number(gsub('S.', '', rmse.files[f]))) %>%
  transform(value = ifelse(is.nan(value), 0, value))

rmse <- rbind(rmse, temp)

} #rds file

rmse <- rmse %>%
  filter(rhat < 1.1) 

mean.rmse <- data.frame(rmse) %>%
  group_by(variable, scenario) %>%
  dplyr::summarize(value = mean(value), .groups = 'keep') %>%
  merge(scenarios, by = 'scenario', all = F) %>%
  group_by(variable, detection, N, survey_freq) %>%
  dplyr::summarize(value = mean(value), .groups = 'keep') %>%
  transform(variable = factor(variable, levels = c('phiP', 'phi1', 'phi2', 'phi3', 'phi4',
                            'phi5', 'phi6', 'phi7', 'phi8', 'phi9', 'phiA',
                            'p1', 'p2', 'pJ', 'pA'))) %>%
  transform(detection = factor(detection, levels = c('Low', 'Med', 'High'))) %>%
  transform(survey_frequency = 'Survey frequency', samp_size = 'Sample size')


#phi
plot_dat <- mean.rmse %>% filter(grepl('phi', variable)) %>%
  transform(variable = factor(variable, levels = c('phiP', 'phi1', 'phi2', 'phi3', 'phi4',
                                                   'phi5', 'phi6', 'phi7', 'phi8', 'phi9', 'phiA'), 
                              labels = phi_pars))

ggplot(plot_dat, aes(x = factor(detection), y = variable, fill = value)) +
  geom_tile(color = 'grey50') +
  xlab('Detection probability') +
  ylab('Parameter') +
  facet_nested(samp_size + N ~ survey_frequency + survey_freq, drop = T, 
             labeller = labeller(survey_freq = freq_labels),
             scales = 'free_x') +
 scale_fill_gradient2(name = "RMSE",
                      mid = "#FFFFFF", low = "#012345", high = "#012345", midpoint = 0) +
  theme_bw() +
  theme(legend.position = 'right',
             strip.background = element_rect(fill = "#EEEEEE", color = "#FFFFFF"),
        legend.text = element_text(size = 8, vjust = 0.5),
        legend.title = element_text(vjust = 0.8, size = 10)) +
    scale_y_discrete("", labels = parse(text = c(eval(expression(levels(plot_dat$variable))))))

#younger ages
ggplot(plot_dat %>% filter(variable %in% c('phi[P]', 'phi[1]', 'phi[2]', 'phi[3]', 'phi[4]',
                                           'phi[5]')), 
       aes(x = factor(detection), y = variable, fill = value)) +
  geom_tile(color = 'grey50') +
  xlab('Detection') +
  ylab('Parameter') +
  facet_nested(samp_size + N ~ survey_frequency + survey_freq, drop = T, 
             labeller = labeller(survey_freq = freq_labels),
             scales = 'free_x') +
 scale_fill_gradient2(name = "RMSE",
                      mid = "#FFFFFF", low = "#012345", high = "#012345", midpoint = 0) +
  theme_bw() +
  theme(legend.position = 'right',
             strip.background = element_rect(fill = "#EEEEEE", color = "#FFFFFF"),
        legend.text = element_text(size = 8, vjust = 0.5),
        legend.title = element_text(vjust = 0.8, size = 10),
        axis.title.x = element_blank()) +
    scale_y_discrete("", labels = parse(text = c(eval(expression(levels(plot_dat$variable))))))

#p
plot_dat <- mean.rmse %>% filter(variable %in% c('p1', 'p2', 'pJ', 'pA')) %>%
  transform(variable = factor(variable, 
                              levels = c('p1', 'p2', 'pJ', 'pA'), 
                              labels = p_pars))

ggplot(plot_dat, aes(x = factor(detection), y = variable, fill = value)) +
  geom_tile(color = 'grey50') +
  xlab('Detection probability') +
  ylab('Parameter') +
  facet_nested(samp_size + N ~ survey_frequency + survey_freq, drop = T, 
             labeller = labeller(survey_freq = freq_labels),
             scales = 'free_x') +
 scale_fill_gradient2(name = "RMSE",
                      mid = "#FFFFFF", low = "#012345", high = "#012345", midpoint = 0) +
  theme_bw() +
  theme(legend.position = 'right',
             strip.background = element_rect(fill = "#EEEEEE", color = "#FFFFFF"),
        legend.text = element_text(size = 8, vjust = 0.5),
        legend.title = element_text(vjust = 0.8, size = 10)) +
    scale_y_discrete("", labels = parse(text = c(eval(expression(levels(plot_dat$variable))))))

```


