---
title: "SSL Future Proof Project Update - Dec 2022"
output: html_document
---

```{r setup, include=FALSE, fig.align = 'center'}

knitr::opts_chunk$set(echo = FALSE, eval = T, message = F, warning = F, fig.align = 'center')

library(here) #v1.0.1
library(dplyr) #v1.0.5
library(nimble) #v0.12.1
library(coda)
library(ggh4x)
library(readr)
library(knitr)
library(Hmisc)
library(stringr)

source(here::here('scripts', 'PlotTheme.R'))

```


### Simulation analysis

The above model was fit using simulated data across 45 scenarios with varying branding cohort sizes (N = 50, 75, 100, 150, 200), branding frequency (annual, biennial, triennial), and detection rates (low, medium, high; low and high = +/- 20% of medium). The only difference between the model in the simulation versus that fit with real data is grouped mean detection probabilities (pups, yearlings, juveniles, adults). No temporal variance was estimated in detection probabilities. Model performance was examined using relative bias, RMSE, and coefficient of variation over both 10- and 20-year study horizons. Resighting effort was assumed to occur annual regardless of branding schedule based on discussions about the likely future reliance on remote cameras. 

The proportion of simulations where all parameters converged (Rhat < 1.1) was high (>95%), though convergence was a bit lower for scenarios with 200 individuals and triennial branding. 

<!-- Table showing the scenario features (branding frequency, detection probability, cohort branding size (N)), the number of simulations per scenario (tot), and the proportion of those simulations where all parameters converged (Rhat < 1.1). Simulations where not all parameters converged were filtered out of the below results.   -->


```{r convergence summary}

sc <- c(1:45)

scenarios <- data.frame(scenario = c(sc),
                        survey_freq = c(rep(1,15), rep(2,15), rep(3,15)),
                        detection = c(rep(c('Low', 'Low', 'Low', 'Low', 'Low',
                                            'Med', 'Med', 'Med', 'Med', 'Med', 
                                            'High', 'High', 'High', 'High', 'High'), 3)),
                        N = c(rep(c(50,75,100,150,200), 9)))
                        
med.files <- list.files(path = here::here('results', 'simpleAge', 'sims', 'prePop'), 
                    pattern = 'median', all.files = FALSE, full.names = F)


converged <- data.frame()
for (f in 1:length(med.files)) {

rhats <- data.frame(rhat = readRDS(file = here::here('results', 'simpleAge','sims', 'prePop',
                                                     med.files[f]))[,'max_rhat']) %>%
  transform(scenario = parse_number(gsub('S.', '', med.files[f]))) %>%
  transform(batch = str_sub(med.files[f], -6,-5)) %>%
  transform(rhat = as.numeric(rhat))

converged <- rbind(converged, rhats)

}

tot_runs <- converged %>%
  filter(!is.na(rhat)) %>%
  group_by(scenario, batch) %>%
  dplyr::summarize(tot = n())

still_need <- tot_runs %>%
  transform(type = ifelse(batch %in% c(10, 14, 18, 22, 26, 30), 'small', 'big')) %>%
  group_by(scenario, type) %>%
  dplyr::summarize(tot = sum(tot)) %>%
  transform(left = 50-tot) %>%
  filter(left != 0)

# prop_conv <- converged %>%
#   filter(rhat<1.10) %>%
#   merge(tot_runs %>% group_by(scenario) %>% dplyr::summarize(tot = sum()), by = c('scenario'), all = T) %>%
#   group_by(scenario) %>%
#   dplyr::summarize(conv = n_distinct(rhat)) %>%
#   transform(prop_conv = round(conv/tot,2)) %>%
#   dplyr::select(scenario, tot, prop_conv) 
# 
# scenarios_tab <- scenarios %>%
#   merge(prop_conv, by = 'scenario') %>%
#   transform(survey_freq = factor(survey_freq, levels = c(1,2,3),
#                                  labels = c('Annual', 'Biennial', 'Triennial')))

# kable(scenarios_tab)


```


#### Relative bias  
Relative bias was generally low for all scenarios, though higher relative bias is evident with triennial branding and low detection probability. Estimates are generally biased high (reds) for pups, yearlings, and juveniles and biased low for age-2 and adults. Patterns in bias for survival estimates are less apparent across levels of detection and sample size than I would have expected, which are more evident in the other model performance metrics below.


```{r results bias}

bias.files <- list.files(path = here::here('results', 'simpleAge', 'sims'), 
                    pattern = 'bias', all.files = FALSE, full.names = F)

rel.bias <- data.frame()
for (f in 1:length(bias.files)) {
temp.bias <- data.frame(readRDS(file = here::here('results', 'simpleAge', 'sims',
                                                  bias.files[f])))
#use gsub on bias files rather than use different list to make sure they match
rhats <- data.frame(rhat = readRDS(file = here::here('results', 'simpleAge', 'sims',
                                                     gsub('bias', 'median', bias.files[f])))[,'max_rhat'])

temp.bias$rhat <- as.numeric(rhats$rhat)
temp <- temp.bias %>% 
  reshape2::melt(id.vars = 'rhat') %>%
  transform(scenario = parse_number(gsub('S.', '', bias.files[f]))) %>%
  transform(value = ifelse(is.nan(value), 0, value))

rel.bias <- rbind(rel.bias, temp)

} #rds file

rel.bias <- rel.bias %>%
  filter(rhat < 1.1) %>%
  transform(length = 10)

mean.rel.bias <- data.frame(rel.bias) %>%
  group_by(variable, scenario) %>%
  dplyr::summarize(value = mean(value), .groups = 'keep') %>%
  merge(scenarios, by = 'scenario', all = F) %>%
  group_by(variable, detection, N, survey_freq) %>%
  dplyr::summarize(value = mean(value)*100, .groups = 'keep') %>%
  transform(variable = factor(variable, levels = c('phiP', 'phi1', 'phi2', 'phiJ', 'phiA',
                            'p1', 'p2', 'pJ', 'pA'))) %>%
  transform(detection = factor(detection, levels = c('Low', 'Med', 'High'))) %>%
  transform(survey_frequency = 'Branding frequency', samp_size = 'Sample size') %>%
  transform(length = 10)

phi_pars <- c('phi[P]','phi[1]', 'phi[2]', 'phi[J]', 'phi[A]')
p_pars <- c('p[1]', 'p[2]', 'p[J]', 'p[A]') 
freq_labels <- c('1' = 'Annual', '2' = 'Biennial', '3' = 'Triennial')

#phi
# plot_dat <- mean.rel.bias %>% filter(grepl('phi', variable)) %>%
#   transform(variable = factor(variable, levels = c('phiP', 'phi1', 'phi2', 'phiJ', 'phiA'), 
#                               labels = phi_pars))
# 
# ggplot(plot_dat, aes(x = factor(detection), y = variable, fill = value)) +
#   geom_tile(color = 'grey50') +
#   xlab('Detection') +
#   ylab('Parameter') +
#   facet_nested(samp_size + N ~ survey_frequency + survey_freq, drop = T, 
#              labeller = labeller(survey_freq = freq_labels),
#              scales = 'free_x') +
#   scale_fill_gradient2(name = "Relative bias (%)",
#                       mid = "#FFFFFF", low = "deepskyblue1", high = "red2", midpoint = 0) +
#   theme_bw() +
#   theme(legend.position = 'right',
#              strip.background = element_rect(fill = "#EEEEEE", color = "#FFFFFF"),
#         legend.text = element_text(size = 8, vjust = 0.5),
#         legend.title = element_text(vjust = 0.8, size = 10)) +
#     scale_y_discrete("", labels = parse(text = c(eval(expression(levels(plot_dat$variable))))))

#p
# plot_dat <- mean.rel.bias %>% filter(variable %in% c('p1', 'p2', 'pJ', 'pA')) %>%
#   transform(variable = factor(variable, 
#                               levels = c('p1', 'p2', 'pJ', 'pA'), 
#                               labels = p_pars))
# 
# ggplot(plot_dat, aes(x = factor(detection), y = variable, fill = value)) +
#   geom_tile(color = 'grey50') +
#   xlab('Detection') +
#   ylab('Parameter') +
#   facet_nested(samp_size + N ~ survey_frequency + survey_freq, drop = T, 
#              labeller = labeller(survey_freq = freq_labels),
#              scales = 'free_x') +
#   scale_fill_gradient2(name = "Relative bias (%)",
#                       mid = "#FFFFFF", low = "deepskyblue1", high = "red2", midpoint = 0) +
#   theme_bw() +
#   theme(legend.position = 'right',
#              strip.background = element_rect(fill = "#EEEEEE", color = "#FFFFFF"),
#         legend.text = element_text(size = 8, vjust = 0.5),
#         legend.title = element_text(vjust = 0.8, size = 10)) +
#     scale_y_discrete("", labels = parse(text = c(eval(expression(levels(plot_dat$variable))))))


```


```{r results bias prepop, fig.width = 6, fig.height = 6}

bias.files <- list.files(path = here::here('results', 'simpleAge', 'sims', 'prePop'), 
                    pattern = 'bias', all.files = FALSE, full.names = F)

rel.bias.20 <- data.frame()
for (f in 1:length(bias.files)) {
temp.bias <- data.frame(readRDS(file = here::here('results', 'simpleAge', 'sims', 'prePop',
                                                  bias.files[f])))
#use gsub on bias files rather than use different list to make sure they match
rhats <- data.frame(rhat = readRDS(file = here::here('results', 'simpleAge', 'sims', 'prePop',
                                                     gsub('bias', 'median', bias.files[f])))[,'max_rhat'])

temp.bias$rhat <- as.numeric(rhats$rhat)
temp <- temp.bias %>% 
  reshape2::melt(id.vars = 'rhat') %>%
  transform(scenario = parse_number(gsub('S.', '', bias.files[f]))) %>%
  transform(value = ifelse(is.nan(value), 0, value))

rel.bias.20 <- rbind(rel.bias.20, temp)

} #rds file

rel.bias.20 <- rel.bias.20 %>%
  filter(rhat < 1.1) #%>%
  # transform(length = 20)

mean.rel.bias.20 <- data.frame(rel.bias.20) %>%
  group_by(variable, scenario) %>%
  dplyr::summarize(value = mean(value), .groups = 'keep') %>%
  merge(scenarios, by = 'scenario', all = F) %>%
  group_by(variable, detection, N, survey_freq) %>%
  dplyr::summarize(value = mean(value)*100, .groups = 'keep') %>%
  transform(variable = factor(variable, levels = c('phiP', 'phi1', 'phi2', 'phiJ', 'phiA',
                            'p1', 'p2', 'pJ', 'pA'))) %>%
  transform(detection = factor(detection, levels = c('Low', 'Med', 'High'))) %>%
  transform(survey_frequency = 'Branding frequency', samp_size = 'Sample size') %>%
  transform(length = 20) %>%
  bind_rows(mean.rel.bias)


phi_pars <- c('phi[P]','phi[1]', 'phi[2]', 'phi[J]', 'phi[A]')
p_pars <- c('p[1]', 'p[2]', 'p[J]', 'p[A]') 
freq_labels <- c('1' = 'Annual', '2' = 'Biennial', '3' = 'Triennial')
occ_labels <- c('10' = '10 years', '20' = '20 years')

#phi
plot_dat <- mean.rel.bias.20 %>% filter(grepl('phi', variable)) %>%
  transform(variable = factor(variable, levels = c('phiP', 'phi1', 'phi2', 'phiJ', 'phiA'), 
                              labels = phi_pars))

#need to add "length" metric" or filter and do dot plot below
ggplot(plot_dat, 
       aes(x = factor(detection), y = variable, fill = value)) +
  geom_tile(color = 'grey50') +
  xlab('Detection') +
  ylab('Parameter') +
  facet_nested(samp_size + N ~ survey_frequency + survey_freq + length, drop = T, 
             labeller = labeller(survey_freq = freq_labels, length = occ_labels),
             scales = 'free_x') +
  scale_fill_gradient2(name = "Relative bias (%)",
                      mid = "#FFFFFF", low = "deepskyblue1", high = "red2", midpoint = 0) +
  theme_bw() +
  theme(legend.position = 'right',
             strip.background = element_rect(fill = "#EEEEEE", color = "#FFFFFF"),
        legend.text = element_text(size = 8, vjust = 0.5),
        legend.title = element_text(vjust = 0.8, size = 10)) +
    scale_y_discrete("", labels = parse(text = c(eval(expression(levels(plot_dat$variable))))))

#p
# plot_dat <- mean.rel.bias %>% filter(variable %in% c('p1', 'p2', 'pJ', 'pA')) %>%
#   transform(variable = factor(variable, 
#                               levels = c('p1', 'p2', 'pJ', 'pA'), 
#                               labels = p_pars))
# 
# ggplot(plot_dat, aes(x = factor(detection), y = variable, fill = value)) +
#   geom_tile(color = 'grey50') +
#   xlab('Detection') +
#   ylab('Parameter') +
#   facet_nested(samp_size + N ~ survey_frequency + survey_freq, drop = T, 
#              labeller = labeller(survey_freq = freq_labels),
#              scales = 'free_x') +
#   scale_fill_gradient2(name = "Relative bias (%)",
#                       mid = "#FFFFFF", low = "deepskyblue1", high = "red2", midpoint = 0) +
#   theme_bw() +
#   theme(legend.position = 'right',
#              strip.background = element_rect(fill = "#EEEEEE", color = "#FFFFFF"),
#         legend.text = element_text(size = 8, vjust = 0.5),
#         legend.title = element_text(vjust = 0.8, size = 10)) +
#     scale_y_discrete("", labels = parse(text = c(eval(expression(levels(plot_dat$variable))))))

# #dot plots
dotplot_dat <- rel.bias.20 %>% transform(length = 20) %>% 
  bind_rows(rel.bias %>% transform(length = 10)) %>%
  merge(scenarios, by = 'scenario', all = F) %>%
  group_by(length, variable, detection, survey_freq, N) %>%
  # dplyr::summarize(value = mean(value),
  #                  lower = quantile(value, probs = 0.025),
  #                  upper = quantile(value, probs = 0.975),
    dplyr::summarize(value = mean(value)*100, .groups = 'keep') %>%
  transform(variable = factor(variable, levels = c('phiP', 'phi1', 'phi2', 'phiJ', 'phiA'))) %>%
  transform(detection = factor(detection, levels = c('Low', 'Med', 'High'))) %>%
  transform(survey_frequency = 'Branding frequency', samp_size = 'Sample size') %>%
  #filter out detection params
  filter(!is.na(variable))


# ggplot(dotplot_dat %>% filter(N %in% c(50, 100, 200)),
#        aes(x = detection, y = value, col = factor(length), group = length)) +
#   geom_point() + geom_line() +
#   geom_hline(aes(yintercept = 0), linetype = 'dotted') +
#   xlab('Detection') + ylab('Relative bias') +
#     # facet_grid(variable~adult_p, labeller = labeller(variable = 'label_parsed'), scales = 'free') +
#   # facet_grid(variable + N ~ survey_freq, label = 'label_parsed', scales = 'free') +
#   
#   facet_nested(variable + samp_size + N ~ survey_frequency + survey_freq, drop = T, 
#              labeller = labeller(survey_freq = freq_labels, length = occ_labels),
#              scales = 'free_x') +
#   
#   theme_bw() +
#   plot_theme(legend.position = 'top',
#              plot.subtitle = element_text(size = 10, hjust = 0.5, vjust = 1)) +
#   scale_color_manual(values = rainbow2[c(5,2)], name = 'Years')

ggplot(dotplot_dat %>% filter(N %in% c(50, 100, 200) & variable %in% c('phiP', 'phiA')),
       aes(x = detection, y = value, col = factor(length), group = length)) +
  geom_point() + geom_line() +
  geom_hline(aes(yintercept = 0), linetype = 'dotted') +
  xlab('Detection') + ylab('Relative bias') +
    # facet_grid(variable~adult_p, labeller = labeller(variable = 'label_parsed'), scales = 'free') +
  # facet_grid(variable + N ~ survey_freq, label = 'label_parsed', scales = 'free') +
  
  facet_nested(variable + samp_size + N ~ survey_frequency + survey_freq, drop = T, 
             labeller = labeller(survey_freq = freq_labels, length = occ_labels),
             scales = 'free_x') +
  
  theme_bw() +
  plot_theme(legend.position = 'top',
             plot.subtitle = element_text(size = 10, hjust = 0.5, vjust = 1)) +
  scale_color_manual(values = rainbow2[c(5,2)], name = 'Years')


```


#### RMSE  
RMSE is a combined measure of bias and variance and is generally very low. Model performance in terms of RMSE follows an expected pattern with lowest RMSE in scenarios with annual branding frequency and higher cohort sizes. RMSE values are generally higher for age-2 and younger, again likely due to their higher variability by design. 


```{r results rmse}

rmse.files <- list.files(path = here::here('results', 'simpleAge', 'sims'), 
                    pattern = 'rmse', all.files = FALSE, full.names = F)

rmse <- data.frame()
for (f in 1:length(rmse.files)) {
temp.rmse <- data.frame(readRDS(file = here::here('results', 'simpleAge', 'sims',
                                                  rmse.files[f])))
#use gsub on rmse files rather than use different list to make sure they match
rhats <- data.frame(rhat = readRDS(file = here::here('results', 'simpleAge', 'sims', 
                                                     gsub('rmse', 'median', rmse.files[f])))[,'max_rhat'])

temp.rmse$rhat <- as.numeric(rhats$rhat)
temp <- temp.rmse %>% 
  reshape2::melt(id.vars = 'rhat') %>%
  transform(scenario = parse_number(gsub('S.', '', rmse.files[f]))) %>%
  transform(value = ifelse(is.nan(value), 0, value))

rmse <- rbind(rmse, temp)

} #rds file

rmse <- rmse %>%
  filter(rhat < 1.1) #%>%
  # transform(length = 10)

mean.rmse <- data.frame(rmse) %>%
  group_by(variable, scenario) %>%
  dplyr::summarize(value = mean(value), .groups = 'keep') %>%
  merge(scenarios, by = 'scenario', all = F) %>%
  group_by(variable, detection, N, survey_freq) %>%
  dplyr::summarize(value = mean(value), .groups = 'keep') %>%
  transform(variable = factor(variable, levels = c('phiP', 'phi1', 'phi2', 'phiJ', 'phiA',
                            'p1', 'p2', 'pJ', 'pA'))) %>%
  transform(detection = factor(detection, levels = c('Low', 'Med', 'High'))) %>%
  transform(survey_frequency = 'Branding frequency', samp_size = 'Sample size') %>%
  transform(length = 10) 


#phi
# plot_dat <- mean.rmse %>% filter(grepl('phi', variable)) %>%
#   transform(variable = factor(variable, levels = c('phiP', 'phi1', 'phi2', 'phiJ', 'phiA'), 
#                               labels = phi_pars))
# 
# ggplot(plot_dat, aes(x = factor(detection), y = variable, fill = value)) +
#   geom_tile(color = 'grey50') +
#   xlab('Detection') +
#   ylab('Parameter') +
#   facet_nested(samp_size + N ~ survey_frequency + survey_freq, drop = T, 
#              labeller = labeller(survey_freq = freq_labels),
#              scales = 'free_x') +
#  scale_fill_gradient2(name = "RMSE",
#                       mid = "#FFFFFF", low = "#012345", high = "#012345", midpoint = 0) +
#   theme_bw() +
#   theme(legend.position = 'right',
#              strip.background = element_rect(fill = "#EEEEEE", color = "#FFFFFF"),
#         legend.text = element_text(size = 8, vjust = 0.5),
#         legend.title = element_text(vjust = 0.8, size = 10)) +
#     scale_y_discrete("", labels = parse(text = c(eval(expression(levels(plot_dat$variable))))))
# 
# #p
# plot_dat <- mean.rmse %>% filter(variable %in% c('p1', 'p2', 'pJ', 'pA')) %>%
#   transform(variable = factor(variable, 
#                               levels = c('p1', 'p2', 'pJ', 'pA'), 
#                               labels = p_pars))
# 
# ggplot(plot_dat, aes(x = factor(detection), y = variable, fill = value)) +
#   geom_tile(color = 'grey50') +
#   xlab('Detection') +
#   ylab('Parameter') +
#   facet_nested(samp_size + N ~ survey_frequency + survey_freq, drop = T, 
#              labeller = labeller(survey_freq = freq_labels),
#              scales = 'free_x') +
#  scale_fill_gradient2(name = "RMSE",
#                       mid = "#FFFFFF", low = "#012345", high = "#012345", midpoint = 0) +
#   theme_bw() +
#   theme(legend.position = 'right',
#              strip.background = element_rect(fill = "#EEEEEE", color = "#FFFFFF"),
#         legend.text = element_text(size = 8, vjust = 0.5),
#         legend.title = element_text(vjust = 0.8, size = 10)) +
#     scale_y_discrete("", labels = parse(text = c(eval(expression(levels(plot_dat$variable))))))

```


```{r results rmse prepop, fig.width = 6, fig.align = 6}

rmse.files <- list.files(path = here::here('results', 'simpleAge', 'sims', 'prePop'), 
                    pattern = 'rmse', all.files = FALSE, full.names = F)

rmse.20 <- data.frame()
for (f in 1:length(rmse.files)) {
temp.rmse <- data.frame(readRDS(file = here::here('results', 'simpleAge', 'sims', 'prePop',
                                                  rmse.files[f])))
#use gsub on rmse files rather than use different list to make sure they match
rhats <- data.frame(rhat = readRDS(file = here::here('results', 'simpleAge', 'sims', 'prePop',
                                                     gsub('rmse', 'median', rmse.files[f])))[,'max_rhat'])

temp.rmse$rhat <- as.numeric(rhats$rhat)
temp <- temp.rmse %>% 
  reshape2::melt(id.vars = 'rhat') %>%
  transform(scenario = parse_number(gsub('S.', '', rmse.files[f]))) %>%
  transform(value = ifelse(is.nan(value), 0, value))

rmse.20 <- rbind(rmse.20, temp)

} #rds file

rmse.20 <- rmse.20 %>%
  filter(rhat < 1.1) #%>%
  # transform(length = 20) 

mean.rmse.20 <- data.frame(rmse.20) %>%
  group_by(variable, scenario) %>%
  dplyr::summarize(value = mean(value), .groups = 'keep') %>%
  merge(scenarios, by = 'scenario', all = F) %>%
  group_by(variable, detection, N, survey_freq) %>%
  dplyr::summarize(value = mean(value), .groups = 'keep') %>%
  transform(variable = factor(variable, levels = c('phiP', 'phi1', 'phi2', 'phiJ', 'phiA',
                            'p1', 'p2', 'pJ', 'pA'))) %>%
  transform(detection = factor(detection, levels = c('Low', 'Med', 'High'))) %>%
  transform(survey_frequency = 'Branding frequency', samp_size = 'Sample size') %>%
  transform(length = 20) %>%
  bind_rows(mean.rmse)


#phi
plot_dat <- mean.rmse.20 %>% filter(grepl('phi', variable)) %>%
  transform(variable = factor(variable, levels = c('phiP', 'phi1', 'phi2', 'phiJ', 'phiA'), 
                              labels = phi_pars))

ggplot(plot_dat, aes(x = factor(detection), y = variable, fill = value)) +
  geom_tile(color = 'grey50') +
  xlab('Detection') +
  ylab('Parameter') +
  facet_nested(samp_size + N ~ survey_frequency + survey_freq + length, drop = T, 
             labeller = labeller(survey_freq = freq_labels, length = occ_labels),
             scales = 'free_x') +
 scale_fill_gradient2(name = "RMSE",
                      mid = "#FFFFFF", low = "#012345", high = "#012345", midpoint = 0) +
  theme_bw() +
  theme(legend.position = 'right',
             strip.background = element_rect(fill = "#EEEEEE", color = "#FFFFFF"),
        legend.text = element_text(size = 8, vjust = 0.5),
        legend.title = element_text(vjust = 0.8, size = 10)) +
    scale_y_discrete("", labels = parse(text = c(eval(expression(levels(plot_dat$variable))))))

#p
# plot_dat <- mean.rmse %>% filter(variable %in% c('p1', 'p2', 'pJ', 'pA')) %>%
#   transform(variable = factor(variable, 
#                               levels = c('p1', 'p2', 'pJ', 'pA'), 
#                               labels = p_pars))
# 
# ggplot(plot_dat, aes(x = factor(detection), y = variable, fill = value)) +
#   geom_tile(color = 'grey50') +
#   xlab('Detection') +
#   ylab('Parameter') +
#   facet_nested(samp_size + N ~ survey_frequency + survey_freq, drop = T, 
#              labeller = labeller(survey_freq = freq_labels),
#              scales = 'free_x') +
#  scale_fill_gradient2(name = "RMSE",
#                       mid = "#FFFFFF", low = "#012345", high = "#012345", midpoint = 0) +
#   theme_bw() +
#   theme(legend.position = 'right',
#              strip.background = element_rect(fill = "#EEEEEE", color = "#FFFFFF"),
#         legend.text = element_text(size = 8, vjust = 0.5),
#         legend.title = element_text(vjust = 0.8, size = 10)) +
#     scale_y_discrete("", labels = parse(text = c(eval(expression(levels(plot_dat$variable))))))


# #dot plots
dotplot_dat <-rmse.20 %>% transform(length = 20) %>% 
  bind_rows(rmse %>% transform(length = 10)) %>%
  merge(scenarios, by = 'scenario', all = F) %>%
  group_by(length, variable, detection, survey_freq, N) %>%
  # dplyr::summarize(value = mean(value),
  #                  lower = quantile(value, probs = 0.025),
  #                  upper = quantile(value, probs = 0.975),
    dplyr::summarize(value = mean(value), .groups = 'keep') %>%
  transform(variable = factor(variable, levels = c('phiP', 'phi1', 'phi2', 'phiJ', 'phiA'))) %>%
  transform(detection = factor(detection, levels = c('Low', 'Med', 'High'))) %>%
  transform(survey_frequency = 'Branding frequency', samp_size = 'Sample size') %>%
  #filter out detection params
  filter(!is.na(variable))

# ggplot(dotplot_dat %>% filter(N %in% c(50, 100, 200)),
#        aes(x = detection, y = value, col = factor(length), group = length)) +
#   geom_point() + geom_line() +
#   geom_hline(aes(yintercept = 0), linetype = 'dotted') +
#   xlab('Detection') + ylab('RMSE') +
#     # facet_grid(variable~adult_p, labeller = labeller(variable = 'label_parsed'), scales = 'free') +
#   # facet_grid(variable + N ~ survey_freq, label = 'label_parsed', scales = 'free') +
#   
#   facet_nested(variable + samp_size + N ~ survey_frequency + survey_freq, drop = T, 
#              labeller = labeller(survey_freq = freq_labels, length = occ_labels),
#              scales = 'free_x') +
#   
#   theme_bw() +
#   plot_theme(legend.position = 'top',
#              plot.subtitle = element_text(size = 10, hjust = 0.5, vjust = 1)) +
#   scale_color_manual(values = rainbow2[c(5,2)], name = 'Years')

# ggplot(dotplot_dat %>% filter(N %in% c(50, 100, 200) & variable %in% c('phiP', 'phiA')),
#        aes(x = detection, y = value, col = factor(length), group = length)) +
#   geom_point() + geom_line() +
#   geom_hline(aes(yintercept = 0), linetype = 'dotted') +
#   xlab('Detection') + ylab('RMSE') +
#     # facet_grid(variable~adult_p, labeller = labeller(variable = 'label_parsed'), scales = 'free') +
#   # facet_grid(variable + N ~ survey_freq, label = 'label_parsed', scales = 'free') +
#   
#   facet_nested(variable + samp_size + N ~ survey_frequency + survey_freq, drop = T, 
#              labeller = labeller(survey_freq = freq_labels, length = occ_labels),
#              scales = 'free_x') +
#   
#   theme_bw() +
#   plot_theme(legend.position = 'top',
#              plot.subtitle = element_text(size = 10, hjust = 0.5, vjust = 1)) +
#   scale_color_manual(values = rainbow2[c(5,2)], name = 'Years')

```


#### Coefficient of variation (CV)  
Lowest CV values are achieved with annual branding and higher cohort sample size. For biennial and triennial branding frequency, estimates for pup survival have higher CV's even with the highest sample sizes (likely because they have a higher simulated standard deviation). Biennial (n = 50) and triennial (all sample sizes) branding frequency does result in CV's above the 0.125 rate that is often mentioned in the MMPA permit applications.


```{r results cv}

cv.files4 <- list.files(path = here::here('results', 'simpleAge', 'sims'), 
                    pattern = 'cv4', all.files = FALSE, full.names = F)

cv.files5 <- list.files(path = here::here('results', 'simpleAge', 'sims'), 
                    pattern = 'cv5', all.files = FALSE, full.names = F)

cv.files6 <- list.files(path = here::here('results', 'simpleAge', 'sims'), 
                    pattern = 'cv6', all.files = FALSE, full.names = F)

cv.files7 <- list.files(path = here::here('results', 'simpleAge', 'sims'), 
                    pattern = 'cv7', all.files = FALSE, full.names = F)

cv.files <- c(cv.files4, cv.files5, cv.files6, cv.files7)

cv <- data.frame()
for (f in 1:length(cv.files)) {
temp.cv <- data.frame(readRDS(file = here::here('results', 'simpleAge', 'sims',
                                                  cv.files[f])))
#use gsub on cv files rather than use different list to make sure they match
rhats <- data.frame(rhat = readRDS(file = here::here('results', 'simpleAge', 'sims', 
                                                     gsub('cv', 'median', cv.files[f])))[,'max_rhat'])

temp.cv$rhat <- as.numeric(rhats$rhat)
temp <- temp.cv %>% 
  reshape2::melt(id.vars = 'rhat') %>%
  transform(scenario = parse_number(gsub('S.', '', cv.files[f]))) %>%
  transform(value = ifelse(is.nan(value), 0, value))

cv <- rbind(cv, temp)

} #rds file

cv <- cv %>%
  filter(rhat < 1.1) 

mean.cv <- data.frame(cv) %>%
  group_by(variable, scenario) %>%
  dplyr::summarize(value = mean(value), .groups = 'keep') %>%
  merge(scenarios, by = 'scenario', all = F) %>%
  group_by(variable, detection, N, survey_freq) %>%
  dplyr::summarize(value = mean(value), .groups = 'keep') %>%
  transform(variable = factor(variable, levels = c('phiP', 'phi1', 'phi2', 'phiJ', 'phiA',
                            'p1', 'p2', 'pJ', 'pA'))) %>%
  transform(detection = factor(detection, levels = c('Low', 'Med', 'High'))) %>%
  transform(survey_frequency = 'Branding frequency', samp_size = 'Sample size') %>%
  transform(length = 10)


#phi
# plot_dat <- mean.cv %>% filter(grepl('phi', variable)) %>%
#   transform(variable = factor(variable, levels = c('phiP', 'phi1', 'phi2', 'phiJ', 'phiA'), 
#                               labels = phi_pars))
# 
# ggplot(plot_dat, aes(x = factor(detection), y = variable, fill = value)) +
#   geom_tile(color = 'grey50') +
#   xlab('Detection') +
#   ylab('Parameter') +
#   facet_nested(samp_size + N ~ survey_frequency + survey_freq, drop = T, 
#              labeller = labeller(survey_freq = freq_labels),
#              scales = 'free_x') +
#  scale_fill_gradient2(name = "cv",
#                       mid = "#FFFFFF", low = "#012345", high = "#012345", midpoint = 0) +
#   theme_bw() +
#   theme(legend.position = 'right',
#              strip.background = element_rect(fill = "#EEEEEE", color = "#FFFFFF"),
#         legend.text = element_text(size = 8, vjust = 0.5),
#         legend.title = element_text(vjust = 0.8, size = 10)) +
#     scale_y_discrete("", labels = parse(text = c(eval(expression(levels(plot_dat$variable))))))

mean.cv.thresh <- data.frame(cv) %>%
  transform(value_cat = value<0.125) %>%
  group_by(variable, scenario) %>%
  dplyr::summarize(value_cat = mean(value_cat), .groups = 'keep') %>%
  merge(scenarios, by = 'scenario', all = F) %>%
  group_by(variable, detection, N, survey_freq) %>%
  dplyr::summarize(value_cat = mean(value_cat), .groups = 'keep') %>%
  transform(variable = factor(variable, levels = c('phiP', 'phi1', 'phi2', 'phiJ', 'phiA'))) %>%
  transform(detection = factor(detection, levels = c('Low', 'Med', 'High'))) %>%
  transform(survey_frequency = 'Branding frequency', samp_size = 'Sample size') %>%
  transform(length = 10)


#phi
# plot_dat <- mean.cv.thresh %>% filter(grepl('phi', variable)) %>%
#   transform(variable = factor(variable, levels = c('phiP', 'phi1', 'phi2', 'phiJ', 'phiA'), 
#                               labels = phi_pars))
# 
# ggplot(plot_dat, aes(x = factor(detection), y = variable, fill = value_cat)) +
#   geom_tile(color = 'grey50') +
#   xlab('Detection') +
#   ylab('Parameter') +
#   facet_nested(samp_size + N ~ survey_frequency + survey_freq, drop = T, 
#              labeller = labeller(survey_freq = freq_labels),
#              scales = 'free_x') +
# scale_fill_gradient2(name = "Probability of cv < 0.125",
#                   mid = "#FFFFFF", high = "deepskyblue1", low = "red2", midpoint = 0.8) +
#   theme_bw() +
#   theme(legend.position = 'right',
#              strip.background = element_rect(fill = "#EEEEEE", color = "#FFFFFF"),
#         legend.text = element_text(size = 8, vjust = 0.5),
#         legend.title = element_text(vjust = 0.8, size = 10)) +
#     scale_y_discrete("", labels = parse(text = c(eval(expression(levels(plot_dat$variable))))))


```


```{r results cv prepop}

cv.files10 <- list.files(path = here::here('results', 'simpleAge', 'sims', 'prePop'), 
                    pattern = 'cv10', all.files = FALSE, full.names = F)

cv.files11 <- list.files(path = here::here('results', 'simpleAge', 'sims', 'prePop'), 
                    pattern = 'cv11', all.files = FALSE, full.names = F)

cv.files14 <- list.files(path = here::here('results', 'simpleAge', 'sims', 'prePop'), 
                    pattern = 'cv14', all.files = FALSE, full.names = F)

cv.files15 <- list.files(path = here::here('results', 'simpleAge', 'sims', 'prePop'), 
                    pattern = 'cv15', all.files = FALSE, full.names = F)

cv.files18 <- list.files(path = here::here('results', 'simpleAge', 'sims', 'prePop'), 
                    pattern = 'cv18', all.files = FALSE, full.names = F)

cv.files19 <- list.files(path = here::here('results', 'simpleAge', 'sims', 'prePop'), 
                    pattern = 'cv19', all.files = FALSE, full.names = F)

cv.files22 <- list.files(path = here::here('results', 'simpleAge', 'sims', 'prePop'),
                         pattern = 'cv22', all.files = FALSE, full.names = F)

cv.files23 <- list.files(path = here::here('results', 'simpleAge', 'sims', 'prePop'), 
                    pattern = 'cv23', all.files = FALSE, full.names = F)

cv.files26 <- list.files(path = here::here('results', 'simpleAge', 'sims', 'prePop'), 
                    pattern = 'cv26', all.files = FALSE, full.names = F)

cv.files27 <- list.files(path = here::here('results', 'simpleAge', 'sims', 'prePop'), 
                    pattern = 'cv27', all.files = FALSE, full.names = F)

cv.files30 <- list.files(path = here::here('results', 'simpleAge', 'sims', 'prePop'), 
                    pattern = 'cv30', all.files = FALSE, full.names = F)

cv.files31 <- list.files(path = here::here('results', 'simpleAge', 'sims', 'prePop'), 
                    pattern = 'cv31', all.files = FALSE, full.names = F)


cv.files <- c(cv.files10, cv.files11, cv.files14, cv.files15, cv.files18, cv.files19,
              cv.files22, cv.files23, cv.files26, cv.files27, cv.files30, cv.files31)

cv.20 <- data.frame()
for (f in 1:length(cv.files)) {
temp.cv <- data.frame(readRDS(file = here::here('results', 'simpleAge', 'sims', 'prePop',
                                                  cv.files[f])))
#use gsub on cv files rather than use different list to make sure they match
rhats <- data.frame(rhat = readRDS(file = here::here('results', 'simpleAge', 'sims', 'prePop',
                                                     gsub('cv', 'median', cv.files[f])))[,'max_rhat'])

temp.cv$rhat <- as.numeric(rhats$rhat)
temp <- temp.cv %>% 
  reshape2::melt(id.vars = 'rhat') %>%
  transform(scenario = parse_number(gsub('S.', '', cv.files[f]))) %>%
  transform(value = ifelse(is.nan(value), 0, value))

cv.20 <- rbind(cv.20, temp)

} #rds file

cv.20 <- cv.20 %>%
  filter(rhat < 1.1) #%>%
  # transform(length = 20) #%>%
  # bind_rows(cv)

mean.cv.20 <- data.frame(cv.20) %>%
  group_by(variable, scenario) %>%
  dplyr::summarize(value = mean(value), .groups = 'keep') %>%
  merge(scenarios, by = 'scenario', all = F) %>%
  group_by(variable, detection, N, survey_freq) %>%
  dplyr::summarize(value = mean(value), .groups = 'keep') %>%
  transform(variable = factor(variable, levels = c('phiP', 'phi1', 'phi2', 'phiJ', 'phiA',
                            'p1', 'p2', 'pJ', 'pA'))) %>%
  transform(detection = factor(detection, levels = c('Low', 'Med', 'High'))) %>%
  transform(survey_frequency = 'Branding frequency', samp_size = 'Sample size') %>%
  transform(length = 20) %>%
  bind_rows(mean.cv)


#phi
plot_dat <- mean.cv.20 %>% filter(grepl('phi', variable)) %>%
  transform(variable = factor(variable, levels = c('phiP', 'phi1', 'phi2', 'phiJ', 'phiA'), 
                              labels = phi_pars))

ggplot(plot_dat, aes(x = factor(detection), y = variable, fill = value)) +
  geom_tile(color = 'grey50') +
  xlab('Detection') +
  ylab('Parameter') +
  facet_nested(samp_size + N ~ survey_frequency + survey_freq + length, drop = T, 
             labeller = labeller(survey_freq = freq_labels, length = occ_labels),
             scales = 'free_x') +
 scale_fill_gradient2(name = "cv",
                      mid = "#FFFFFF", low = "#012345", high = "#012345", midpoint = 0) +
  theme_bw() +
  theme(legend.position = 'right',
             strip.background = element_rect(fill = "#EEEEEE", color = "#FFFFFF"),
        legend.text = element_text(size = 8, vjust = 0.5),
        legend.title = element_text(vjust = 0.8, size = 10)) +
    scale_y_discrete("", labels = parse(text = c(eval(expression(levels(plot_dat$variable))))))

#threshold approach
mean.cv.thresh.20 <- data.frame(cv %>% transform(length = 10)) %>% 
  bind_rows(cv.20 %>% transform(length = 20)) %>%
  transform(value_cat = value<0.125) %>%
  group_by(variable, scenario, length) %>%
  dplyr::summarize(value_cat = mean(value_cat), .groups = 'keep') %>%
  merge(scenarios, by = 'scenario', all = F) %>%
  group_by(variable, detection, N, survey_freq, length) %>%
  dplyr::summarize(value_cat = mean(value_cat), .groups = 'keep') %>%
  transform(variable = factor(variable, levels = c('phiP', 'phi1', 'phi2', 'phiJ', 'phiA'))) %>%
  transform(detection = factor(detection, levels = c('Low', 'Med', 'High'))) %>%
  transform(survey_frequency = 'Branding frequency', samp_size = 'Sample size')


#phi
plot_dat <- mean.cv.thresh.20 %>% filter(grepl('phi', variable)) %>%
  transform(variable = factor(variable, levels = c('phiP', 'phi1', 'phi2', 'phiJ', 'phiA'), 
                              labels = phi_pars))

ggplot(plot_dat, aes(x = factor(detection), y = variable, fill = value_cat)) +
  geom_tile(color = 'grey50') +
  xlab('Detection') +
  ylab('Parameter') +
  facet_nested(samp_size + N ~ survey_frequency + survey_freq + length, drop = T, 
             labeller = labeller(survey_freq = freq_labels, length = occ_labels),
             scales = 'free_x') +
scale_fill_gradient2(name = "Probability of cv < 0.125",
                  mid = "#FFFFFF", high = "deepskyblue1", low = "red2", midpoint = 0.8) +
  theme_bw() +
  theme(legend.position = 'right',
             strip.background = element_rect(fill = "#EEEEEE", color = "#FFFFFF"),
        legend.text = element_text(size = 8, vjust = 0.5),
        legend.title = element_text(vjust = 0.8, size = 10)) +
    scale_y_discrete("", labels = parse(text = c(eval(expression(levels(plot_dat$variable))))))


```

#### Detecting changes in survival  
To examine the ability to detect a change in survival from a "baseline", I generated annual survival estimates using a logit normal distribution with a mean rate derived by multiplying the empirical mean by an age-based change threshold and a standard deviation from the empirical estimates. I then examined the probability of whether the credible intervals for each survival rate excluded the original "baseline" rate at 5, 10, 15, and 20 years into the study. I examined the ability to detect a "small" decrease (15%, 10%, 5%, 2.5%) and a "big" decrease (20%, 15%, 10%, 5%) in survival for pups, age-1:2, juveniles, and adults, respectively. 

The first figure shows the probability of detecting a big decrease at 15 versus 20 years, which are fairly similar. The second figure shows the probability of detecting a small versus big decrease at 20 years. This figure generally illustrates the trade-offs between branding cohort size and branding frequency for the probability of detecting small (harder) and big (easier) changes in survival. For example, detecting small changes in survival over a 20-year period is fairly unlikely without annual branding. Detecting bigger changes in survival with a biennial branding frequency become more probable with higher sample sizes. 


```{r results threshold}

library(stringr)

cri.scenarios <- c('dec_little', 'dec_big', 'inc_little', 'inc_big')
cri.scenarios <- c('dec_little', 'dec_big')
append <- c(5,10)

cri.all <- data.frame()

for (a in 1:length(append)) { #files for period length

for (j in 1:length(cri.scenarios)) {
  
cri.files <- list.files(path = here::here('results', 'simpleAge', 'sims'), 
                    pattern = paste0(cri.scenarios[j],append[a]), all.files = FALSE, full.names = F)

cri <- data.frame()
for (f in 1:length(cri.files)) {
temp.cri <- data.frame(readRDS(file = here::here('results', 'simpleAge', 'sims',
                                                  cri.files[f])))
#use gsub on cri files rather than use different list to make sure they match
rhats <- data.frame(rhat = readRDS(file = here::here('results', 'simpleAge', 'sims',
                                                     gsub('median_', 'median', gsub(paste0(cri.scenarios[j], append[a]), 'median', cri.files[f]))))[,'max_rhat'])

temp.cri$rhat <- as.numeric(rhats$rhat)
temp <- temp.cri %>% 
  reshape2::melt(id.vars = 'rhat') %>%
  transform(scenario = parse_number(gsub('S.', '', cri.files[f]))) %>%
  transform(run = str_sub(cri.files[f], -5,-5)) %>%
  transform(length = append[a]) %>%
  # transform(length = ifelse(grepl('little5', cri.files[f]) | grepl('big5', cri.files[f]), 5, 10)) %>%
  transform(threshold = cri.scenarios[j])

cri <- rbind(cri, temp)

} #appending length of year period

cri.all <- rbind(cri, cri.all)

} #rds file
} #j


mean.cri <- data.frame(cri.all) %>%
    merge(scenarios, by = 'scenario', all = F) %>%
  #cri only correct for runs 5 and greater
  # filter(run > 6) %>%
  group_by(threshold, variable, detection, N, survey_freq, length) %>%
  dplyr::summarize(value = mean(value), .groups = 'keep') %>%
  transform(variable = factor(variable, levels = c('phiP', 'phi1', 'phi2', 'phiJ', 'phiA',
                            'p1', 'p2', 'pJ', 'pA'))) %>%
  transform(detection = factor(detection, levels = c('Low', 'Med', 'High'))) %>%
  transform(survey_frequency = 'Branding frequency', samp_size = 'Sample size') 


# plot_dat <- mean.cri %>% filter(grepl('phi', variable)) %>%
#   transform(variable = factor(variable, levels = c('phiP', 'phi1', 'phi2', 'phiJ', 'phiA'), 
#                               labels = phi_pars)) %>%
#   transform(value_cat = ifelse(value < 0.5, "<0.5",
#                                ifelse(value >=0.5 & value < 0.8, '0.5-0.8', '>0.8'))) %>%
#   transform(value_cat = factor(value_cat, levels = c('<0.5', '0.5-0.8', '>0.8'))) %>%
#   transform(threshold = factor(threshold, levels = c('dec_little', 'dec_big', 'inc_little', 'inc_big'),
#                                labels = c('Small decrease', 'Big decrease', 'Small increase', 'Big increase')))


#continuous color scale
#decrease in survival with 5 and 10 yrs of data
# ggplot(plot_dat %>% filter(threshold %in% c('Big decrease')), 
#        aes(x = factor(detection), y = variable, fill = value)) +
#         # aes(x = factor(detection), y = variable, fill = value_cat)) +
#   geom_tile(color = 'grey50') +
#   xlab('Detection') +
#   ylab('Parameter') +
#   facet_nested(samp_size + N ~ survey_frequency + survey_freq + length, drop = T, 
#              labeller = labeller(survey_freq = freq_labels),
#              scales = 'free_x') +
# scale_fill_gradient2(name = "Probability",
#                   mid = "#FFFFFF", high = "deepskyblue1", low = "red2", midpoint = 0.8) +
#   theme_bw() +
#   theme(legend.position = 'top',
#              strip.background = element_rect(fill = "#EEEEEE", color = "#FFFFFF")) +
#     scale_y_discrete("", labels = parse(text = c(eval(expression(levels(plot_dat$variable)))))) 

#decrease in survival with 10 yrs of data
# ggplot(plot_dat %>% filter(threshold %in% c('Small decrease', 'Big decrease') & length == 10), 
#        aes(x = factor(detection), y = variable, fill = value)) +
#         # aes(x = factor(detection), y = variable, fill = value_cat)) +
#   geom_tile(color = 'grey50') +
#   xlab('Detection') +
#   ylab('Parameter') +
#   facet_nested(samp_size + N ~ survey_frequency + survey_freq + threshold, drop = T, 
#              labeller = labeller(survey_freq = freq_labels),
#              scales = 'free_x') +
# scale_fill_gradient2(name = "Probability",
#                   mid = "#FFFFFF", high = "deepskyblue1", low = "red2", midpoint = 0.8) +
#   theme_bw() +
#   theme(legend.position = 'top',
#              strip.background = element_rect(fill = "#EEEEEE", color = "#FFFFFF")) +
#     scale_y_discrete("", labels = parse(text = c(eval(expression(levels(plot_dat$variable)))))) 

#categorical colors
# ggplot(plot_dat %>% filter(threshold %in% c('Small decrease', 'Big decrease') & length == 10),
#        # aes(x = factor(detection), y = variable, fill = value)) +
#         aes(x = factor(detection), y = variable, fill = value_cat)) +
#   geom_tile(color = 'grey50') +
#   xlab('Detection') +
#   ylab('Parameter') +
#   facet_nested(samp_size + N ~ survey_frequency + survey_freq + threshold, drop = T,
#              labeller = labeller(survey_freq = freq_labels),
#              scales = 'free_x') +
#   theme_bw() +
#   theme(legend.position = 'top',
#              strip.background = element_rect(fill = "#EEEEEE", color = "#FFFFFF")) +
#     scale_y_discrete("", labels = parse(text = c(eval(expression(levels(plot_dat$variable)))))) +
#   scale_fill_manual(name = "Probability of detecting change in survival",
#                     values = c('firebrick3', 'orange', 'steelblue2'))
# 
# ggplot(plot_dat %>% filter(threshold %in% c('Big decrease')),
#        # aes(x = factor(detection), y = variable, fill = value)) +
#         aes(x = factor(detection), y = variable, fill = value_cat)) +
#   geom_tile(color = 'grey50') +
#   xlab('Detection') +
#   ylab('Parameter') +
#   facet_nested(samp_size + N ~ survey_frequency + survey_freq + length, drop = T,
#              labeller = labeller(survey_freq = freq_labels),
#              scales = 'free_x') +
#   theme_bw() +
#   theme(legend.position = 'top',
#              strip.background = element_rect(fill = "#EEEEEE", color = "#FFFFFF")) +
#     scale_y_discrete("", labels = parse(text = c(eval(expression(levels(plot_dat$variable)))))) +
#   scale_fill_manual(name = "Probability of detecting change in survival",
#                     values = c('firebrick3', 'orange', 'steelblue2'))
# 
# #dot plots
# dotplot_dat <- cri.all %>%
#   merge(scenarios, by = 'scenario', all = F) %>%
#   group_by(threshold, length, variable, detection, survey_freq, N) %>%
#   # dplyr::summarize(value = mean(value),
#   #                  lower = quantile(value, probs = 0.025),
#   #                  upper = quantile(value, probs = 0.975),
#     dplyr::summarize(value = mean(value), .groups = 'keep') %>%
#   transform(variable = factor(variable, levels = c('phiP', 'phi1', 'phi2', 'phiJ', 'phiA'))) %>%
#   transform(detection = factor(detection, levels = c('Low', 'Med', 'High'))) %>%
#   transform(survey_frequency = 'Branding frequency', samp_size = 'Sample size') %>%
#   transform(threshold = factor(threshold, levels = c('dec_little', 'dec_big', 'inc_little', 'inc_big'),
#                                labels = c('Small decrease', 'Big decrease', 'Small increase', 'Big increase')))
# 
# 
# ggplot(plot_dat %>% filter(threshold %in% c('Small decrease', 'Big decrease')),
#        aes(x = N, y = value, col = length, group = length)) +
#   geom_point() + geom_line() +
#   geom_hline(aes(yintercept = 0), linetype = 'dotted') +
#   xlab('Sample size') + ylab('Probability of detecting change') +
#     # facet_grid(variable~adult_p, labeller = labeller(variable = 'label_parsed'), scales = 'free') +
#   facet_grid(variable + detection ~ survey_freq + threshold, label = 'label_parsed', scales = 'free') +
#   theme_bw() +
#   plot_theme(legend.position = 'top',
#              plot.subtitle = element_text(size = 10, hjust = 0.5, vjust = 1)) +
#   scale_color_manual(values = rainbow2[c(5,2)], name = 'Threshold')



```


```{r results threshold prepop, fig.width = 7, fig.height = 6}

library(stringr)

cri.scenarios <- c('dec_little', 'dec_big', 'inc_little', 'inc_big')
cri.scenarios <- c('dec_little', 'dec_big')
append <- c(5,10)

cri.all.20 <- data.frame()

for (a in 1:length(append)) { #files for period length

for (j in 1:length(cri.scenarios)) {
  
cri.files <- list.files(path = here::here('results', 'simpleAge', 'sims', 'prePop'), 
                    pattern = paste0(cri.scenarios[j],append[a]), all.files = FALSE, full.names = F)

cri <- data.frame()
for (f in 1:length(cri.files)) {
temp.cri <- data.frame(readRDS(file = here::here('results', 'simpleAge', 'sims', 'prePop',
                                                  cri.files[f])))
#use gsub on cri files rather than use different list to make sure they match
rhats <- data.frame(rhat = readRDS(file = here::here('results', 'simpleAge', 'sims', 'prePop', 
                                                     gsub('median_', 'median', gsub(paste0(cri.scenarios[j], append[a]), 'median', cri.files[f]))))[,'max_rhat'])

temp.cri$rhat <- as.numeric(rhats$rhat)
temp <- temp.cri %>% 
  reshape2::melt(id.vars = 'rhat') %>%
  transform(scenario = parse_number(gsub('S.', '', cri.files[f]))) %>%
  transform(run = str_sub(cri.files[f], -6,-5)) %>%
  transform(length = append[a]+10) %>%
  # transform(length = ifelse(grepl('little5', cri.files[f]) | grepl('big5', cri.files[f]), 5, 10)) %>%
  transform(threshold = cri.scenarios[j])

cri <- rbind(cri, temp)

} #appending length of year period

cri.all.20 <- rbind(cri, cri.all.20)

} #rds file
} #j

mean.cri.20 <- data.frame(cri.all.20) %>%
  #are these just all the unfinished/unfilled sim dataframe rows? 
  filter(!is.na(value)) %>%
    merge(scenarios, by = 'scenario', all = F) %>%
  group_by(threshold, variable, detection, N, survey_freq, length) %>%
  dplyr::summarize(value = mean(value), .groups = 'keep') %>%
  transform(variable = factor(variable, levels = c('phiP', 'phi1', 'phi2', 'phiJ', 'phiA'))) %>%
  transform(detection = factor(detection, levels = c('Low', 'Med', 'High'))) %>%
  transform(survey_frequency = 'Branding frequency', samp_size = 'Sample size') %>%
  bind_rows(mean.cri)


plot_dat <- mean.cri.20 %>% filter(grepl('phi', variable)) %>%
  transform(variable = factor(variable, levels = c('phiP', 'phi1', 'phi2', 'phiJ', 'phiA'), 
                              labels = phi_pars)) %>%
  transform(value_cat = ifelse(value < 0.5, "<0.5",
                               ifelse(value >=0.5 & value < 0.8, '0.5-0.8', '>0.8'))) %>%
  transform(value_cat = factor(value_cat, levels = c('<0.5', '0.5-0.8', '>0.8'))) %>%
  transform(threshold = factor(threshold, levels = c('dec_little', 'dec_big', 'inc_little', 'inc_big'),
                               labels = c('Small decrease', 'Big decrease', 'Small increase', 'Big increase')))


#continuous color scale
#decrease in survival with 15 and 20 yrs of data
ggplot(plot_dat %>% filter(threshold %in% c('Big decrease')), 
       aes(x = factor(detection), y = variable, fill = value)) +
        # aes(x = factor(detection), y = variable, fill = value_cat)) +
  geom_tile(color = 'grey50') +
  xlab('Detection') +
  ylab('Parameter') +
  facet_nested(samp_size + N ~ survey_frequency + survey_freq + length, drop = T, 
             labeller = labeller(survey_freq = freq_labels),
             scales = 'free_x') +
scale_fill_gradient2(name = "Probability of detecting change",
                  mid = "#FFFFFF", high = "deepskyblue1", low = "red2", midpoint = 0.8) +
  theme_bw() +
  theme(legend.position = 'top',
             strip.background = element_rect(fill = "#EEEEEE", color = "#FFFFFF")) +
    scale_y_discrete("", labels = parse(text = c(eval(expression(levels(plot_dat$variable)))))) 

#decrease in survival with 20 yrs of data
# ggplot(plot_dat %>% filter(threshold %in% c('Small decrease', 'Big decrease') & length == 20),
#        aes(x = factor(detection), y = variable, fill = value)) +
#         # aes(x = factor(detection), y = variable, fill = value_cat)) +
#   geom_tile(color = 'grey50') +
#   xlab('Detection') +
#   ylab('Parameter') +
#   facet_nested(samp_size + N ~ survey_frequency + survey_freq + threshold, drop = T,
#              labeller = labeller(survey_freq = freq_labels),
#              scales = 'free_x') +
# scale_fill_gradient2(name = "Probability of detecting change",
#                   mid = "#FFFFFF", high = "deepskyblue1", low = "red2", midpoint = 0.8) +
#   theme_bw() +
#   theme(legend.position = 'top',
#              strip.background = element_rect(fill = "#EEEEEE", color = "#FFFFFF")) +
#     scale_y_discrete("", labels = parse(text = c(eval(expression(levels(plot_dat$variable))))))

# #more facets?
# ggplot(plot_dat,
#        aes(x = factor(detection), y = variable, fill = value)) +
#         # aes(x = factor(detection), y = variable, fill = value_cat)) +
#   geom_tile(color = 'grey50') +
#   xlab('Detection') +
#   ylab('Parameter') +
#   facet_nested(samp_size + N + length ~ survey_frequency + survey_freq + threshold, drop = T,
#              labeller = labeller(survey_freq = freq_labels),
#              scales = 'free_x') +
# scale_fill_gradient2(name = "Probability of detecting change",
#                   mid = "#FFFFFF", high = "deepskyblue1", low = "red2", midpoint = 0.8) +
#   theme_bw() +
#   theme(legend.position = 'top',
#              strip.background = element_rect(fill = "#EEEEEE", color = "#FFFFFF")) +
#     scale_y_discrete("", labels = parse(text = c(eval(expression(levels(plot_dat$variable))))))

#categorical colors
# ggplot(plot_dat %>% filter(threshold %in% c('Small decrease', 'Big decrease') & length == 10),
#        # aes(x = factor(detection), y = variable, fill = value)) +
#         aes(x = factor(detection), y = variable, fill = value_cat)) +
#   geom_tile(color = 'grey50') +
#   xlab('Detection') +
#   ylab('Parameter') +
#   facet_nested(samp_size + N ~ survey_frequency + survey_freq + threshold, drop = T,
#              labeller = labeller(survey_freq = freq_labels),
#              scales = 'free_x') +
#   theme_bw() +
#   theme(legend.position = 'top',
#              strip.background = element_rect(fill = "#EEEEEE", color = "#FFFFFF")) +
#     scale_y_discrete("", labels = parse(text = c(eval(expression(levels(plot_dat$variable)))))) +
#   scale_fill_manual(name = "Probability of detecting change in survival",
#                     values = c('firebrick3', 'orange', 'steelblue2'))
# 
# ggplot(plot_dat %>% filter(threshold %in% c('Big decrease')),
#        # aes(x = factor(detection), y = variable, fill = value)) +
#         aes(x = factor(detection), y = variable, fill = value_cat)) +
#   geom_tile(color = 'grey50') +
#   xlab('Detection') +
#   ylab('Parameter') +
#   facet_nested(samp_size + N ~ survey_frequency + survey_freq + length, drop = T,
#              labeller = labeller(survey_freq = freq_labels),
#              scales = 'free_x') +
#   theme_bw() +
#   theme(legend.position = 'top',
#              strip.background = element_rect(fill = "#EEEEEE", color = "#FFFFFF")) +
#     scale_y_discrete("", labels = parse(text = c(eval(expression(levels(plot_dat$variable)))))) +
#   scale_fill_manual(name = "Probability of detecting change in survival",
#                     values = c('firebrick3', 'orange', 'steelblue2'))
# 
# #dot plots
# dotplot_dat <- cri.all.20 %>%
#   merge(scenarios, by = 'scenario', all = F) %>%
#   group_by(threshold, length, variable, detection, survey_freq, N) %>%
#   # dplyr::summarize(value = mean(value),
#   #                  lower = quantile(value, probs = 0.025),
#   #                  upper = quantile(value, probs = 0.975),
#     dplyr::summarize(value = mean(value), .groups = 'keep') %>%
#   transform(variable = factor(variable, levels = c('phiP', 'phi1', 'phi2', 'phiJ', 'phiA'))) %>%
#   transform(detection = factor(detection, levels = c('Low', 'Med', 'High'))) %>%
#   transform(survey_frequency = 'Branding frequency', samp_size = 'Sample size') %>%
#   transform(threshold = factor(threshold, levels = c('dec_little', 'dec_big', 'inc_little', 'inc_big'),
#                                labels = c('Small decrease', 'Big decrease', 'Small increase', 'Big increase')))
# 
# 
# ggplot(plot_dat %>% filter(threshold %in% c('Small decrease', 'Big decrease')),
#        aes(x = N, y = value, col = length, group = length)) +
#   geom_point() + geom_line() +
#   geom_hline(aes(yintercept = 0), linetype = 'dotted') +
#   xlab('Sample size') + ylab('Probability of detecting change') +
#     # facet_grid(variable~adult_p, labeller = labeller(variable = 'label_parsed'), scales = 'free') +
#   facet_grid(variable + detection ~ survey_freq + threshold, label = 'label_parsed', scales = 'free') +
#   theme_bw() +
#   plot_theme(legend.position = 'top',
#              plot.subtitle = element_text(size = 10, hjust = 0.5, vjust = 1)) +
#   scale_color_manual(values = rainbow2[c(5,2)], name = 'Threshold')



```


```{r more facets, fig.width = 7.5, fig.height = 9}

plot_dat <- mean.cri.20 %>% filter(grepl('phi', variable)) %>%
  transform(variable = factor(variable, levels = c('phiP', 'phi1', 'phi2', 'phiJ', 'phiA'), 
                              labels = phi_pars)) %>%
  transform(value_cat = ifelse(value < 0.5, "<0.5",
                               ifelse(value >=0.5 & value < 0.8, '0.5-0.8', '>0.8'))) %>%
  transform(value_cat = factor(value_cat, levels = c('<0.5', '0.5-0.8', '>0.8'))) %>%
  transform(threshold = factor(threshold, levels = c('dec_little', 'dec_big', 'inc_little', 'inc_big'),
                               labels = c('Small decrease', 'Big decrease', 'Small increase', 'Big increase')))

#more facets
ggplot(plot_dat,
       aes(x = factor(detection), y = variable, fill = value)) +
        # aes(x = factor(detection), y = variable, fill = value_cat)) +
  geom_tile(color = 'grey50') +
  xlab('Detection') +
  ylab('Parameter') +
  facet_nested(samp_size + N + length ~ survey_frequency + survey_freq + threshold, drop = T,
             labeller = labeller(survey_freq = freq_labels),
             scales = 'free_x') +
scale_fill_gradient2(name = "Probability of detecting change",
                  mid = "#FFFFFF", high = "deepskyblue1", low = "red2", midpoint = 0.8) +
  theme_bw() +
  theme(legend.position = 'top',
             strip.background = element_rect(fill = "#EEEEEE", color = "#FFFFFF")) +
    scale_y_discrete("", labels = parse(text = c(eval(expression(levels(plot_dat$variable))))))

```



### Empirical demographic and detection rate estimates (eDPS)  

Survival rates were estimated for ages P, 1, 2, juveniles (ages 3-4) adults (ages 5+), with fixed effects estimating the effect of pup mass and the differences between sex and natal region on survival over 19 years of data. The sample sizes for natal region were imbalanced, with more samples from Pyramid Rock (*n* = 1395) than South Seal Rock (*n* = 290). Fixed effects of sex were estimated for each age group while those of natal region were estimated for pups, yearlings, shared for juveniles (ages 3-4) and adults (ages 7+). Temporal variance was estimated using a random effect of year with a penalized complexity prior, with the random effect estimated for pups, yearlings, and pooled for juveniles (ages 3-4) and adults (ages 5+) jointly for both sex and natal region (i.e., trends were not estimated for each sex nor region but for age groups in the eDPS as a whole). Detection rates were estimated for each age as above with temporal variance constructed in the same fashion, but detection was assumed to be constant across sexes, natal regions, and resight location.  

#### Survival across sex
```{r phi by sex}

out <- readRDS(here::here('results', 'simpleAge', 'out_ME.RDS'))

all_pars <- colnames(out[[1]])
noZ <- all_pars[which(!grepl('z', all_pars))]

posts <- out[,noZ]
outmat <- as.matrix(posts)

post_sum <- data.frame(
  med = apply(outmat, 2, function(x) quantile(x, probs = 0.5, na.rm = T, names = F)),
  lower = apply(outmat, 2, function(x) quantile(x, probs = 0.025, na.rm = T, names = F)),
  upper = apply(outmat, 2, function(x) quantile(x, probs = 0.975, na.rm = T, names = F)))
post_sum$variable <- row.names(post_sum)

#mean rates
sex.pars <- all_pars[which(grepl('b.sex', all_pars))]
mu.pars <- all_pars[which(grepl('mu.', all_pars))]

pars <- c(sex.pars, mu.pars)

mu.vals <- post_sum %>%
  filter(variable %in% mu.pars) %>%
  transform(age = gsub('mu.', '', variable)) 

b.sex <- post_sum %>% filter(grepl('b.sex', variable) & med != 0) %>%
  transform(age = ifelse(grepl('sexP', variable), 'P', 
                         ifelse(grepl('sex1', variable), '1',
                                ifelse(grepl('sex2', variable), '2', 
                                       ifelse(grepl('sex3', variable), '3', 
                                              ifelse(grepl('sexJ', variable), 'J', 'A'))))))

#Female back transformed values
mean.phi.sex.vals <- mu.vals %>%
  merge(b.sex, by = 'age', suffixes = c('', '.sex'), all = T) %>%
  #fill b.sex for older age intercepts
  # transform(med.sex = ifelse(is.na(med.sex), b.sex[b.sex$age == 'A', 'med'], med.sex),
  #           lower.sex = ifelse(is.na(lower.sex), b.sex[b.sex$age == 'A', 'lower'], lower.sex),
  #           upper.sex = ifelse(is.na(upper.sex), b.sex[b.sex$age == 'A', 'upper'], upper.sex)) %>%
  transform(med = 1/(1+exp(-(med + med.sex))),
            lower = 1/(1+exp(-(lower + lower.sex))),
            upper = 1/(1+exp(-(upper + upper.sex)))) %>%
  transform(sex = 'Female') %>%
  dplyr::select(c(med, lower, upper, variable, age, sex)) %>%
  bind_rows(
    post_sum %>%
  filter(grepl('int.phi', variable)) %>%
  transform(age = gsub('int.phi', '', variable), sex = 'Male')) %>%
  transform(age = factor(age, levels = c('P', '1', '2', '3', 'J', 'A'),
                         labels = c('P', '1', '2', '3', 'J', 'A'))) 

ggplot(mean.phi.sex.vals, aes(age, med, col = sex, group = sex)) +
  geom_point(position = position_dodge(0.5)) +
  geom_errorbar(aes(ymin = lower, ymax = upper), width = 0.25, position = position_dodge(0.5)) +
  xlab('') + ylab(expression(paste('Survival probability ',  '(', phi, ')'))) +
  ggtitle('') +
  plot_theme(legend.position = 'top', panel.border = element_rect(fill = NA),
             plot.title = element_text(hjust = 0.5)) +
  guides(color = guide_legend("", nrow = 1, byrow = T)) +
  scale_color_manual(values = rainbow2[c(2,5)], name = '') 

```

#### Survival by site  
```{r phi by site}

#mean rates
site.pars <- all_pars[which(grepl('b.site', all_pars))]
mu.pars <- all_pars[which(grepl('mu.', all_pars))]

pars <- c(site.pars, mu.pars)

mu.vals <- post_sum %>%
  filter(variable %in% mu.pars) %>%
  transform(age = gsub('mu.', '', variable)) 

b.site <- post_sum %>% filter(grepl('b.site', variable) & med != 0) %>%
  transform(age = ifelse(grepl('siteP', variable), 'P', 
                         ifelse(grepl('site1', variable), '1',
                                ifelse(grepl('site2', variable), '2', 
                                       ifelse(grepl('siteJ', variable), 'J', 'A')))))

#male back transformed values
mean.phi.vals <- mu.vals %>%
  merge(b.site, by = 'age', suffixes = c('', '.site'), all.x = T) %>%
  #fill b.sex for older age intercepts
  transform(med.site = ifelse(is.na(med.site) & age %in% c(3), b.site[b.site$age == 'J', 'med'], med.site),
            lower.site = ifelse(is.na(lower.site) & age %in% c(3), b.site[b.site$age == 'J', 'lower'], lower.site),
            upper.site = ifelse(is.na(upper.site) & age %in% c(3), b.site[b.site$age == 'J', 'upper'], upper.site)) %>%
  transform(med = 1/(1+exp(-(med + med.site))),
            lower = 1/(1+exp(-(lower + lower.site))),
            upper = 1/(1+exp(-(upper + upper.site)))) %>%
  transform(site = 'South Seal Rock') %>%
  dplyr::select(c(med, lower, upper, variable, age, site)) %>%
  bind_rows(
    post_sum %>%
  filter(grepl('int.phi', variable)) %>%
  transform(age = gsub('int.phi', '', variable), site = 'Pyramid Rock')) %>%
  transform(age = factor(age, levels = c('P', '1', '2', '3', 'J', 'A'),
                         labels = c('P', '1', '2', '3', 'J', 'A'))) 

ggplot(mean.phi.vals, aes(age, med, col = site, group = site)) +
  geom_point(position = position_dodge(0.5)) +
  geom_errorbar(aes(ymin = lower, ymax = upper), width = 0.25, position = position_dodge(0.5)) +
  xlab('') + ylab(expression(paste('Survival probability ',  '(', phi, ')'))) +
  ggtitle('') +
  plot_theme(legend.position = 'top', panel.border = element_rect(fill = NA),
             plot.title = element_text(hjust = 0.5)) +
  guides(color = guide_legend("", nrow = 1, byrow = T)) +
  scale_color_manual(values = rainbow2[c(2,5)], name = '') 


```

#### Survival over time

```{r temporal variance}


#temporal variance
phi.vals <- post_sum %>%
  filter(grepl('.prob', variable)) %>%
    transform(year = as.numeric(as.character(gsub('.+\\[([0-9]+)\\].*$', '\\1',
                                                variable)))) %>%
  #use when dataframe for phi.prob rather than vector
  # transform(year = as.numeric(as.character(gsub('.+\\[([0-9]+), ([0-9]+)\\].*$', '\\2', 
  #                                               variable)))) %>% 
  # transform(region = gsub('.+\\[([0-9]+), ([0-9]+)\\].*$', '\\1', variable)) %>%
  transform(variable = sub("\\[.*", "", variable)) %>%
  transform(age = gsub('.prob', '', variable)) %>%
  transform(age = gsub('phi', '', age)) %>%
  filter(age %in% c('P', '1', '2', '3', '4', 'A')) %>%
  transform(age_cat = factor(age, levels = c('P', '1', '2', '3', '4', 'A'))) %>%
  transform(age = factor(age, levels = c('P', '1', '2', '3', '4', 'A'),
                         labels = c('P', '1', '2', '3', 'J', 'A'))) %>%
  merge(mean.phi.vals %>% filter(site == 'Pyramid Rock'), by = c('age'), suffixes = c('', '.mean'))

#temporal variance with mean rates; intercept represents female, so can't show males
ggplot(phi.vals,
       aes(year, med)) +
  geom_errorbar(aes(x = year, ymin=lower, ymax=upper), position = position_dodge(0.5),
                width = 0.5, show.legend = F) +
  geom_point(size = 0.8, position = position_dodge(0.5)) +
  geom_line(size = 0.7, position = position_dodge(0.5)) +
  geom_line(aes(year, med.mean), col = 'grey20', linetype = 'dotted') +
  xlab('') + ylab(expression(paste('Survival probability ',  '(', phi, ')'))) +
  ggtitle('') +
    facet_wrap(. ~ age) +
  plot_theme(legend.position = 'top', panel.border = element_rect(fill = NA),
             plot.title = element_text(hjust = 0.5)) +
  guides(color = guide_legend("", nrow = 1, byrow = T)) +
  scale_color_manual(values = rainbow2[c(2,3,5,6,7,8)], name = '') +
  scale_x_continuous(breaks = c(seq(1, 19, by = 2)), labels = c(seq(2001, 2019, by = 2)))


```

### Detection probability
```{r detection}

#mean detection probabilities
p.vals <- post_sum %>%
  filter(!grepl('phi', variable) & grepl('int.p', variable)) %>%
  transform(age = gsub('int.p', '', variable)) %>%
  transform(age = factor(age, levels = c('1', '2', '3', 'J', 'A'),
                         labels = c('1', '2', '3', 'J', 'A')))

ggplot(p.vals, aes(age, med)) +
  geom_point() +
  geom_errorbar(aes(ymin = lower, ymax = upper), width = 0.15) +
  xlab('') + ylab('Detection probability') +
  ggtitle('') +
  plot_theme(legend.position = 'top', panel.border = element_rect(fill = NA),
             plot.title = element_text(hjust = 0.5)) +
  guides(color = guide_legend("", nrow = 1, byrow = T)) +
  scale_color_manual(values = rainbow2[c(2,3,5,6,7,8)], name = '') 

#temporal variance
p.vals <- post_sum %>%
  filter(grepl('.prob', variable)) %>%
  filter(!grepl('phi', variable)) %>%
  transform(year = as.numeric(as.character(gsub('.+\\[([0-9]+)\\].*$', '\\1', 
                                                variable)))) %>%
  transform(variable = sub("\\[.*", "", variable)) %>%
  transform(age = gsub('.prob', '', variable)) %>%
  transform(age = gsub('p', '', age)) %>%
  transform(age = factor(age, levels = c('1', '2', '3', '4', 'A'),
                         labels = c('1', '2', '3', 'J', 'A'))) %>%
  filter(!is.na(age))

ggplot(p.vals, aes(year, med)) +
  geom_errorbar(aes(x = year, ymin=lower, ymax=upper), position = position_dodge(0.5), 
                width = 0.5, show.legend = F) +
  geom_point(size = 0.8, position = position_dodge(0.5)) + 
  geom_line(size = 0.7, position = position_dodge(0.5)) +
  xlab('') + ylab(expression(paste('Detection probability '))) +
  ggtitle('') +
    facet_wrap(. ~ age) +
  # facet_grid(age ~ region) +
  plot_theme(legend.position = 'top', panel.border = element_rect(fill = NA),
             plot.title = element_text(hjust = 0.5)) +
  guides(color = guide_legend("", nrow = 1, byrow = T)) +
  scale_color_manual(values = rainbow2[c(2,3,5,6,7,8)], name = '') +
  scale_x_continuous(breaks = c(seq(1, 19, by = 2)), labels = c(seq(2001, 2019, by = 2)))

```


